---
title: "Starting with Bathy for Shelves"
output:
  pdf_document: default
  html_notebook: default
---

```{r setup}
library(raster)
library(rgdal)
library(sf)
library(ncdf4)
#library(rmapshaper)
library(tidyverse)
library(diptest)
library(moments)
library(viridis) #colors
library(data.table)
library(hydroTSM) #hypsometric curves
library(gridExtra)
library(ggfortify)
library(rgeos)
library(ggridges) #allows for gradient in density plot
library(psych) #summary statistics
library(gridExtra)

etopo_shelf_df <- readRDS("raw_data/etopo_shelf_df.rds")
#bring in bathy for shelf regions

#LMEs
LME_spdf <- readOGR("raw_data/LME66/LMEs66.shp")

#keep Antarctica and the arctic
LME_poles <- LME_spdf[LME_spdf$LME_NUMBER == 61 | LME_spdf$LME_NUMBER == 64,]


#get rid of Antarctica and the arctic
LME_spdf <- LME_spdf[LME_spdf$LME_NUMBER != 61 & LME_spdf$LME_NUMBER != 64,]


#equal area projection
eaproj <- CRS("+proj=laea")
LME_spdf.EA <- spTransform(LME_spdf, eaproj)


```


Let's just start with one LME to get started
```{r LME_baltic}
LME_baltic <- LME_spdf[1,]
plot(LME_baltic)

center <- gCentroid(LME_baltic)
center@coords[2]
```


Make bathy layer into raster (I should see if this is different if I use equal area projection (To Do))
```{r rasterize}
proj <- crs(LME_spdf)
etopo_shelf_raster <- rasterFromXYZ(etopo_shelf_df, crs = proj)


#make figure of shelves
  #png("etopo_shelf_bathy.png", height=5, width=7, units="in", res=300)
  #plot(etopo_shelf_raster)
  #dev.off()

```

Depth distribution
```{r depth distribution}
mean(etopo_shelf_raster$z)
hist(etopo_shelf_raster$z)

cellStats(etopo_shelf_raster, stat = 'mean')
hist(etopo_shelf_raster)
```
Depth distribution for Antarctica and Arctic specifically
```{r Antarctic & Arctic}
#crop bathy layer to LME subset
bathy_arctic_extent <- crop(etopo_shelf_raster, extent(LME_poles[2,]))

#which areas of raster fall within borders?
bathy_arctic_mask <- mask(bathy_arctic_extent, LME_poles[2,])

#convert to df for plotting

  # First, to a SpatialPointsDataFrame
  bathy_arctic_pts <- rasterToPoints(bathy_arctic_mask, spatial = TRUE)
  # Then to a 'conventional' dataframe
  bathy_arctic_dt  <- data.table(data.frame(bathy_arctic_pts))

#keep depth positive  
bathy_arctic_dt[,depth := -(z)]

#fortify to plot shapefile on top
#LME_arctic_df <- fortify(LME_arctic) (don't need because indstead I make polygon to surround raster)

#To get a polygon that surrounds cells that are not NA

  # make all values the same.
  bathy_arctic_mask_1s <- reclassify(bathy_arctic_mask, cbind(-Inf, Inf, 1))
  
  # convert to polygons (you need to have package 'rgeos' installed for this to work)
  bathy_arctic_polygon_1s <- rasterToPolygons(bathy_arctic_mask_1s, dissolve=TRUE) #single polygon
  bathy_arctic_polygon <- rasterToPolygons(bathy_arctic_mask, dissolve=TRUE) #one polygon for each depth profile
  
  LME_reduced_arctic_df <- fortify(bathy_arctic_polygon)

ggplot(bathy_arctic_dt) +
  geom_tile(aes(x = x, y = y, fill = depth)) +
  scale_fill_viridis(direction = -1, limits = c(0,2000)) +
  theme_classic() +
  labs(x = paste0("Longitude ","\u00B0","E"), y = paste0("Latitude ","\u00B0","N"), fill = "Depth (m)") +
  guides(fill = guide_colorbar(reverse = T)) +
  geom_path(data = LME_reduced_arctic_df, aes(x = long, y = lat, group = group), fill = NA, color = "black")
```


Clip raster to each individual polygon

```{r clip raster to polygon}
#crop bathy layer to LME subset
bathy_baltic_extent <- crop(etopo_shelf_raster, extent(LME_baltic))

#which areas of raster fall within borders?
bathy_baltic_mask <- mask(bathy_baltic_extent, LME_baltic)

#convert to df for plotting

  # First, to a SpatialPointsDataFrame
  bathy_baltic_pts <- rasterToPoints(bathy_baltic_mask, spatial = TRUE)
  # Then to a 'conventional' dataframe
  bathy_baltic_dt  <- data.table(data.frame(bathy_baltic_pts))

#keep depth positive  
bathy_baltic_dt[,depth := -(z)]

#fortify to plot shapefile on top
#LME_baltic_df <- fortify(LME_baltic) (don't need because indstead I make polygon to surround raster)

#To get a polygon that surrounds cells that are not NA

  # make all values the same.
  bathy_baltic_mask_1s <- reclassify(bathy_baltic_mask, cbind(-Inf, Inf, 1))
  
  # convert to polygons (you need to have package 'rgeos' installed for this to work)
  bathy_baltic_polygon_1s <- rasterToPolygons(bathy_baltic_mask_1s, dissolve=TRUE) #single polygon
  bathy_baltic_polygon <- rasterToPolygons(bathy_baltic_mask, dissolve=TRUE) #one polygon for each depth profile
  
  LME_reduced_baltic_df <- fortify(bathy_baltic_polygon)

ggplot(bathy_baltic_dt) +
  geom_tile(aes(x = x, y = y, fill = depth)) +
  scale_fill_viridis(direction = -1, limits = c(0,2000)) +
  theme_classic() +
labs(x = paste0("Longitude ","\u00B0","E"), y = paste0("Latitude ","\u00B0","N"), fill = "Depth (m)") +
  guides(fill = guide_colorbar(reverse = T)) +
  geom_path(data = LME_reduced_baltic_df, aes(x = long, y = lat, group = group), fill = NA, color = "black")

summary(bathy_baltic_dt)

#how to get area of raster

  #get sizes of all cells in raster [km2]
  cell_size<-area(bathy_baltic_mask, na.rm=TRUE, weights=FALSE)
  #delete NAs from vector of all raster cells
  ##NAs lie outside of the rastered region, can thus be omitted
  cell_size<-cell_size[!is.na(cell_size)]
  #compute area [km2] of all cells in geo_raster
  bathy_raster_area <- sum(cell_size)
  
#Also, try using raster::area on polygon

#reproject
#The Lambert azimuthal equal-area projection is a particular mapping from a sphere to a disk. It accurately represents area in all regions of the sphere, but it does not accurately represent angles.
  equalareaprojection<- crs("+proj=laea")

  bathy_baltic_polygon_1s.EA <- spTransform(bathy_baltic_polygon_1s, equalareaprojection)
  bathy_baltic_polygon.EA <- spTransform(bathy_baltic_polygon, equalareaprojection)
  
    #get sizes of all polygons in shapefile (m^2)
  polygon_area <- area(bathy_baltic_polygon.EA, dissolve = T)

  polygon_area_kms <- polygon_area/1e6

```

Will have to be creative on how to map LMEs that cross 180 longitude, ex. Aleutian Islands
```{r high longitude examples}
LME_datatable <- LME_spdf@data
LME_aleutian <- LME_spdf[LME_spdf$LME_NUMBER == 65,]

#crop bathy layer to LME subset
bathy_aleutian_extent <- crop(etopo_shelf_raster, extent(LME_aleutian))

#which areas of raster fall within borders?
bathy_aleutian_mask <- mask(bathy_aleutian_extent, LME_aleutian)

#convert to df for plotting

  # First, to a SpatialPointsDataFrame
  bathy_aleutian_pts <- rasterToPoints(bathy_aleutian_mask, spatial = TRUE)
  # Then to a 'conventional' dataframe
  bathy_aleutian_dt  <- data.table(data.frame(bathy_aleutian_pts))

#keep depth positive  
bathy_aleutian_dt[,depth := -(z)]

bathy_aleutian_dt$x_shift <- ifelse(bathy_aleutian_dt$x >0, bathy_aleutian_dt$x - 360, bathy_aleutian_dt$x)


#To get a polygon that surrounds cells that are not NA

  # make all values the same.
  r <- reclassify(bathy_aleutian_mask, cbind(-Inf, Inf, 1))
  
  # convert to polygons (you need to have package 'rgeos' installed for this to work)
  pp <- rasterToPolygons(r, dissolve=TRUE)
  
  LME_reduced_aleutian_df <- fortify(pp)
  
  
LME_reduced_aleutian_df$long_shift <- ifelse(LME_reduced_aleutian_df$long >0, LME_reduced_aleutian_df$long - 360, LME_reduced_aleutian_df$long)
  

ggplot(bathy_aleutian_dt) +
  geom_tile(aes(x = x_shift, y = y, fill = depth)) +
  scale_fill_viridis(direction = -1, limits = c(0,2000)) +
  theme_classic() +
  labs(x = paste0("Longitude ","\u00B0","E"), y = paste0("Latitude ","\u00B0","N"), fill = "Depth (m)") +
  guides(fill = guide_colorbar(reverse = T)) +
  geom_path(data = LME_reduced_aleutian_df, aes(x = long_shift, y = lat, group = group), fill = NA, color = "black")

```

How does area vary with depth
```{r hyposometric curves}
#list of values within raster
bathy_baltic_depth_list <- getValues(bathy_baltic_mask)
bathy_baltic_depth_list <- bathy_baltic_depth_list[!is.na(bathy_baltic_depth_list)]

bathy_baltic_depth_list_pos <- -1*bathy_baltic_depth_list

bathy_baltic_SGDF <- as(bathy_baltic_mask, 'SpatialGridDataFrame') # as spatial grid data frame
hypsometric(bathy_baltic_SGDF)

#dimensions of raster?
dim(bathy_baltic_mask)

#total number of cells?
cells <- dim(bathy_baltic_mask)[1]*dim(bathy_baltic_mask)[2]


#cells of each depth value in data table
bathy_baltic_bydepth <- data.table(freq(bathy_baltic_mask))

#NA cells
na_cell_count <- bathy_baltic_bydepth[is.na(value),]$count

#filled cells

cells_filled <- cells-na_cell_count

bathy_baltic_bydepth_complete <- bathy_baltic_bydepth[!is.na(value)]

#calculate percent of total area in polygon
bathy_baltic_bydepth_complete[,percent_area := count/sum(count)]


bathy_baltic_bydepth_complete[,area := percent_area*bathy_raster_area][,depth := -value]



(plot <- ggplot(data = bathy_baltic_bydepth_complete, aes(x = depth, y = area, fill = depth)) +
  geom_col(width = 10) +
  theme_classic() +
  labs(x = "Depth (m)", y = "Area (km^2)", fill = "Depth (m)") +
  scale_y_continuous(expand = c(0, 0)) +
  scale_x_continuous(expand = c(0,0)) +
  scale_fill_gradientn(colors = rev(viridis(10)), limits = c(0, 2000)) +
  guides(position = "bottom", fill = guide_colorbar(reverse = T)))
  

ggsave(plot, filename = "Baltic_Sea.png")
```


How to classify different LMES? Dip (Modality) & Skew
 "distributions with a dip value >0.01 and with significant (p<0.05) deviations from unimodality to the hourglass classification. 
 
 For distributions with a dip value ≤0.01, we assigned those with a Type-I skewness ≥0.5 to pyramid, 
 
 those with skewness ≤ −0.5 to inverse pyramid, 
 
 and the remainder to diamond, representing those with approximately normal distributions. 
 
 We chose skew cutoffs of 0.5 and −0.5 to capture right- and left-skewed distributions, respectively, and to bound distributions approximating symmetry." from Elsen and Tingley, but I will go with 1 to -1
 
Modality: In mathematics, unimodality means possessing a unique mode. More generally, unimodality means there is only a single highest value, somehow defined, of some mathematical object. (Wikipedia)

Skew: In probability theory and statistics, skewness is a measure of the asymmetry of the probability distribution of a real-valued random variable about its mean. The skewness value can be positive, zero, negative, or undefined. (Wikipedia)

Kolmogorov–Smirnov test for uniformity

null = uniform distribution
goal = reject the null (p<0.05)


```{r uniformity test}
plot(ecdf(bathy_baltic_depth_list_pos))
curve(punif(x, min(bathy_baltic_depth_list_pos), max(bathy_baltic_depth_list_pos)), add=TRUE, col="red")
ks.test <- ks.test(bathy_baltic_depth_list_pos, "punif", min(bathy_baltic_depth_list_pos), max(bathy_baltic_depth_list_pos))

ks.test$p.value

```

"The dip test measures multimodality in a sample by the maximum difference, over all sample points, between the empirical distribution function, and the unimodal distribution function that minimizes that maximum difference."


 Dip Test for Modality
```{r dip test for unimodality}
#Computes Hartigans’ dip test statistic for testing unimodality, and additionally the modal interval.

diptest <- dip.test(bathy_baltic_depth_list_pos, simulate.p.value = TRUE, B = 2000)
p.value <- diptest$p.value

```


Skew

* If skewness is less than -1 or greater than 1, the distribution is highly skewed.**
* If skewness is between -1 and -0.5 or between 0.5 and 1, the distribution is moderately skewed.
* If skewness is between -0.5 and 0.5, the distribution is approximately symmetric.
```{r skew}
skew <- skewness(bathy_baltic_depth_list_pos, na.rm = T)

```


Mean depth value
```{r mean depth}
mean <- mean(bathy_baltic_depth_list_pos)

mean
```

Possible groupings:

Shallow dominant: High positive skew, unimodal
Mid-Dominant: Relatively normal (low skew, unimodal)
Deep-Dominant: High negative skew, unimodal
Uniform: reject the null for Kolmogorov–Smirnov test for uniformity
Hourglass: Bimodality

Text to add to plot:
 - class
 - skew
 - dip
 - mean
 - bottleneck

```{r full plot}

plot +
  annotate("text", x = 300, y = 4300, label =  paste0("Kolmogorov-Smirnov P-Value: ", signif(ks.test$p.value,2))) +
  annotate("text", x = 300, y = 4000, label = paste0("Dip Test: ", signif(diptest$statistic,2), " P-value = ", signif(p.value))) +
  annotate("text", x = 300, y = 3700, label = paste0("Skew: ", round(skew,2))) +
  annotate("text", x = 300, y = 3400, label = paste0("Mean: ",round(mean,2), " m")) +
  geom_vline(xintercept = mean, color = "red")
```
--------
Now we have the methods down, so let's make a loop!!

```{r loop for caculations and classifications}

#plot map, plot hypsometric curve, plot histogram

#reorder LME_spdf by LME # 
LME_spdf_ordered <- LME_spdf[order(LME_spdf$LME_NUMBER),]


LME_bathy_statistics <- as.data.table(matrix(nrow = nrow(LME_spdf)))
LME_bathy_statistics[, lme_name := as.factor(V1)][, lme_number := as.numeric(V1)][,lme_area := as.numeric(V1)][,ks_test_pvalue := as.numeric(V1)][, dip_test := as.numeric(V1)][, dip_p_value := as.numeric(V1)][, skew := as.numeric(V1)][, mean_depth := as.numeric(V1)][, max_depth := as.numeric(V1)][, median_depth := as.numeric(V1)][, center_latitude := as.numeric(V1)]

LME_bathy_statistics[, V1 := NULL]


lme_maps <- vector("list", length = nrow(LME_spdf))
lme_densityplots <- vector("list", length = nrow(LME_spdf))
lme_areaplots <- vector("list", length = nrow(LME_spdf))

for (i in 1:nrow(LME_bathy_statistics)) {
  LME_single <- LME_spdf_ordered[i,]
  LME_bathy_statistics[i, "lme_name"] <- LME_single$LME_NAME
  LME_bathy_statistics[i, "lme_number"] <- LME_single$LME_NUMBER
  
  #extract centroid latitude value
  centroid <- gCentroid(LME_single)
  LME_bathy_statistics[i, "center_latitude"] <- centroid@coords[2]
  
  #clip raster to LME
  LME_extent <- crop(etopo_shelf_raster, extent(LME_single))
  
  #which areas of raster fall within borders?
  LME_mask <- mask(LME_extent, LME_single)
  
  #conver to df for plotting
    # First, to a SpatialPointsDataFrame
    LME_points <- rasterToPoints(LME_mask, spatial = TRUE)
    
  #project points to help with better plotting
 #   LME_points <- spTransform(LME_points, CRS())
    
    # Then to a 'conventional' dataframe
    LME_dt  <- data.table(data.frame(LME_points))
  
  #keep depth positive  
  LME_dt[,depth := -(z)]
  
  LME_dt$x_shift <- ifelse(
    max(LME_dt$x)-min(LME_dt$x) > 200 & LME_dt$x >0,
    LME_dt$x - 360,
    LME_dt$x)
  
  #size of raster  
  #get sizes of all cells in raster [km2]
    cell_size<-area(LME_mask, na.rm=TRUE, weights=FALSE)
    #delete NAs from vector of all raster cells
    ##NAs lie outside of the rastered region, can thus be omitted
    cell_size<-cell_size[!is.na(cell_size)]
    #compute area [km2] of all cells in geo_raster
    bathy_raster_area<-length(cell_size)*median(cell_size)
    
    LME_bathy_statistics[i, "lme_area"] <- bathy_raster_area
    
  #To get a polygon that surrounds cells that are not NA
  
    # make all values the same.
    r <- reclassify(LME_mask, cbind(-Inf, Inf, 1))
    
    # convert to polygons (you need to have package 'rgeos' installed for this to work)
    pp <- rasterToPolygons(r, dissolve=TRUE) #longest step
    
    LME_reduced_df <- fortify(pp)
    
    LME_reduced_df$long_shift <- ifelse(
    max(LME_reduced_df$long)-min(LME_reduced_df$long) > 200 & LME_reduced_df$long >0,
    LME_reduced_df$long - 360,
    LME_reduced_df$long)
    
    depth_palette <- c(
  `yellow`        = "E3E247",
  `green`      = "46AF78",
  `blue`       = "472F7B",
  `darkpurple`     = "280330")
  
  lme_maps[[i]] <- ggplot(LME_dt,aes(x = x_shift, y = y, fill = depth)) +
    geom_tile() +
    scale_fill_gradientn(
  colours=c("#f98f3e","#f9be3e","#E3E247", "#46AF78", "#3B5088","#592f83","#420d54", "#420d54"),
  breaks = c(min(LME_dt$depth),
                    100,
                    200,
                    300,
                    400,
                     500,
                     500,
                     max(LME_dt$depth))) +
    theme_classic() +
      labs(x = paste0("Longitude ","\u00B0","E"), y = paste0("Latitude ","\u00B0","N"), fill = "Depth (m)") +
    guides(fill = guide_colorbar(reverse = T)) +
  #  geom_path(data = LME_reduced_df, aes(x = long_shift, y = lat, group = group), color = "black") +
    labs(title = paste0("LME ", LME_bathy_statistics[i,lme_number], "/n", LME_bathy_statistics[i,lme_name])) + #change direction of  / if running on annotate (PC versus mac problem)
    theme(plot.title = element_text(hjust=0.5, size = 11, face = "bold"))
  
  
    #list of values within raster
  LME_bathy_depth_list <- getValues(LME_mask)
  LME_bathy_depth_list <- LME_bathy_depth_list[!is.na(LME_bathy_depth_list)]
  
  LME_bathy_depth_list_pos <- -1*LME_bathy_depth_list
  
  #assign specific values to each grid cell
  cell_size_cell_value <- data.table(cell_size = cell_size, depth = LME_bathy_depth_list_pos)
  
  #sum area across cells with same depth
  cell_size_cell_value[,sum_area := sum(cell_size), by = depth]

  #reduce to frequency table
  freq_table_raster <- unique(cell_size_cell_value, by = c("depth", "sum_area"))

  freq_table_raster[,cell_size := NULL] #remove cell size column
  
  #calculate percent of total area in polygon
  #generate vector of values for modality calculation with diptest package
  freq_table_raster[,prop_area :=  sum_area/sum(sum_area)] #calculate proportional area

  freq_table_raster[,area_rounddown := round(prop_area * 10000000, 0)] #check, do I actually need this many sig figs?

  #empty vector
  LME_data_vector <- vector()

  #populate vector
    for (i in 1:nrow(freq_table_raster)){
    add <- rep(freq_table_raster[i,depth], freq_table_raster[i,area_rounddown])
    LME_data_vector <- append(LME_data_vector, add, after = length(LME_data_vector))
    
    }
  
  
  #plot
  lme_areaplots[[i]] <- ggplot(data = freq_table_raster, aes(x = depth, y = sum_area, fill = depth)) +
    geom_col(width = 5) +
#    geom_smooth(method = "gam", se = F, color = "black", size = 1) +
    theme_classic() +
    labs(x = "Depth (m)", y = expression(paste("Area (", km^{2},")")), fill = "Depth (m)") +
    scale_y_continuous(expand = c(0, 0)) +
    scale_x_continuous(expand = c(0,0)) +
    scale_fill_gradientn(
  colours=c("#f98f3e","#f9be3e","#E3E247", "#46AF78", "#3B5088","#592f83","#420d54", "#420d54"),
  breaks = c(min(LME_bathy_bydepth_complete$depth),
                    100,
                    200,
                    300,
                    400,
                     500,
                     500,
                     max(LME_bathy_bydepth_complete$depth))) +
    guides(position = "bottom", fill = guide_colorbar(reverse = T))
  
  #uniformity
  ks_test <- ks.test(LME_bathy_depth_list_pos, "punif", min(LME_bathy_depth_list_pos), max(LME_bathy_depth_list_pos))
  LME_bathy_statistics[i,"ks_test_pvalue"] <- ks_test$p.value
  
  #unimodality
  #Computes Hartigans’ dip test statistic for testing unimodality, and additionally the modal interval.
  
   diptest <- dip.test(LME_data_vector, simulate.p.value = TRUE, B = 2000)
#this takes a while, probably more than 30 minutes
  LME_bathy_statistics[i,"dip_test"] <- diptest$statistic[1]
  p.value <- diptest$p.value
  LME_bathy_statistics[i,"dip_p_value"] <- p.value
  
  skew <- skewness(LME_data_vector, na.rm = T)
  LME_bathy_statistics[i,"skew"] <- skew
  
  mean <- mean(LME_data_vector)
  LME_bathy_statistics[i,"mean_depth"] <- mean
  
  max_depth <- max(LME_data_vector)
  LME_bathy_statistics[i, "max_depth"] <- max_depth
  
  median_depth <- median(LME_data_vector)
  LME_bathy_statistics[i, "median_depth"] <- median_depth
  
  max_area <- max(LME_bathy_bydepth_complete$area)
 
    print(paste(i, nrow(LME_bathy_statistics), sep = "/"))
  

}

saveRDS(LME_bathy_statistics, file = "LME_bathy_statistics.rds")
saveRDS(lme_densityplots, file = "lme_densityplots.rds")
saveRDS(lme_areaplots, file = "lme_areaplots.rds")
saveRDS(lme_maps, file = "lme_maps.rds")

#plots to pdf
ggsave("arranged_maps.pdf", marrangeGrob(lme_maps, ncol=2, nrow = 2))
#ggsave("arranged_densityplots.pdf", marrangeGrob(grobs = lme_densityplots, ncol=2, nrow = 3))
ggsave("arranged_areaplots.pdf", marrangeGrob(grobs = lme_areaplots, ncol=2, nrow = 3))
```
Associate LME with region

- keep in mind, LME 1 (1&6) and 54 (1&6) and 65 (1&6) appear in multiple

```{r LME to region}
west_pac <- c(35,	36,	37,	39,	40,	41,	42,	46,	47,	48,	49,	50,	51,	52,	53, 56,	57)
east_pac <- c(2, 3, 4, 11, 13, 55)
west_atl <- c(5, 6, 7, 8, 9, 12, 14, 15, 16, 17, 18, 63, 66)
east_atl <- c(19, 20, 21, 22, 23, 24, 25, 26, 27, 28, 29, 58, 59, 60, 62)
west_ind <- c(30, 31, 32, 33)
east_ind <- c(34, 38, 43, 44, 45)
east_west_pac <- c(1, 54, 65)
east_pac_west_atl <- c(61)

#OR larger groupings (includes LME's left out of earlier analyses because disconnected, aka Hawaii etc.)
pacific <- c(35,	36,	37,	39,	40,	41,	42,	46,	47,	48,	49,	50,	51,	52,	53, 56,	57, 2, 3, 4, 11, 13, 55, 1, 54, 65,10)
atlantic <- c(5, 6, 7, 8, 9, 12, 14, 15, 16, 17, 18, 63, 66, 19, 20, 21, 22, 23, 24, 25, 26, 27, 28, 29, 58, 59, 60, 62 )
indian <- c(30, 31, 32, 33, 34, 38, 43, 44, 45)
antarctica <- c(61) #excluding
arctic <- c(64) #excluding



#LME_bathy_statistics[,region := ifelse(lme_number %in% west_pac, "west_pac",ifelse(lme_number %in% east_pac, #"east_pac", ifelse(lme_number %in% west_atl, "west_atl", ifelse(lme_number %in% west_ind, "west_ind", #ifelse(lme_number %in% east_atl, "east_atl", ifelse(lme_number %in% east_ind, "east_ind", ifelse(lme_number #%in% east_west_pac, "east_west_pac", ifelse(lme_number %in% east_pac_west_atl, "east_pac_west_atl", #"none"))))))))]

LME_bathy_statistics[,region := ifelse(lme_number %in% pacific, "pacific", ifelse(lme_number %in% atlantic, "atlantic", ifelse(lme_number %in% indian, "indian", ifelse(lme_number %in% antarctica, "antarctica", "arctic"))))]

```


Add classifications to data frame

* If skewness is less than -1 or greater than 1, the distribution is highly skewed.
* If skewness is between -1 and -0.5 or between 0.5 and 1, the distribution is moderately skewed.
* If skewness is between -0.5 and 0.5, the distribution is approximately symmetric.

distributions with a dip value >0.01 and with significant (p<0.05) deviations from unimodality to the hourglass classification. 

For distributions with a dip value ≤0.01, we assigned those with a Type-I skewness ≥0.5 to shallow dominant, those with skewness ≤ −0.5 to deep dominant, and the remainder to diamond, representing those with approximately normal distributions. This is what Elsen and Tingley did. Now, I'm going to play around with skew cutoff of 1 rather than 0.5.

If pvalue for uniformity test is >0.05 --> uniform

```{r expanded dataframe}
LME_bathy_statistics[, skew_descriptor := ifelse(abs(skew) > 1, "high_skew", ifelse(abs(skew) < 0.5, "symmetric", "mod_skew"))]

LME_bathy_statistics[, hourglass := ifelse(dip_test > 0.01, ifelse(dip_p_value < 0.05, T, F), F)]

LME_bathy_statistics[, deep_dominant := ifelse(dip_test <= 0.01, 
                                               ifelse(skew <= -1, T, F), F)]

LME_bathy_statistics[, shallow_dominant := ifelse(dip_test <= 0.01, 
                                                  ifelse(skew >= 1, T, F), F)]

LME_bathy_statistics[, mid_dominant := ifelse(dip_test <= 0.01,
                                          ifelse(abs(skew) <1, T, F), F)]

#negative kurtosis (no tails)
LME_bathy_statistics[, uniform := ifelse(ks_test_pvalue >0.05,T,F)]

```

Are all regions labeled by type?
```{r labeling check}
LME_bathy_statistics[,type := ifelse(uniform == T, "Uniform",
                                     ifelse(hourglass == T, "Multimodal",
                                        ifelse(deep_dominant == T, "Deep Dominant",
                                           ifelse(shallow_dominant == T, "Shallow Dominant",
                                                  ifelse(mid_dominant == T, "Mid Dominant", "none")))))][,lme_name2 := lme_name]

write.csv(LME_bathy_statistics, "LME_bathy_statistics.csv")
saveRDS(LME_bathy_statistics, file = "LME_bathy_statistics.rds")

#if already loaded
#LME_bathy_statistics <- readRDS(file = "LME_bathy_statistics.RData") 
```

Now that we have classified distributions, let's make final plots
```{r plots with classification and statistics}

lme_areaplots_classified <- vector("list", length = nrow(LME_bathy_statistics))

for (i in 1:nrow(LME_bathy_statistics)) {
  
  #get max y value to set label location
  ggp <- ggplot_build(lme_areaplots[[i]])
max_y <- ggp$layout$panel_scales_y[[1]]$range$range[[2]]  #y range
  
#full plot
  lme_areaplots_classified[[i]] <- lme_areaplots[[i]] +
    geom_vline(xintercept = LME_bathy_statistics[i,mean_depth], color = "red") +
    annotate("text", x=Inf, y = max_y-max_y/20, hjust=1, label = paste0("LME ", LME_bathy_statistics[i,lme_number], ":\n ", LME_bathy_statistics[i,lme_name]), fontface = "bold", size = 2) +
        annotate("text", x=Inf, y = max_y-3*max_y/20, hjust=1, label = paste0("Classification: ", LME_bathy_statistics[i,type]), size = 2) +
    annotate("text", x=Inf, y = max_y-4*max_y/20, hjust=1, label = paste0("Dip Test: ", signif(LME_bathy_statistics[i,dip_test],2), " P-value = ", signif(LME_bathy_statistics[i,dip_p_value], 2)), size = 2) +
    annotate("text", x=Inf, y = max_y-5*max_y/20, hjust=1, label = paste0("Skew: ", round(LME_bathy_statistics[i,skew],2)), size = 2) +
    annotate("text", x=Inf, y = max_y-6*max_y/20, hjust=1, label = paste0("Mean: ",round(LME_bathy_statistics[i,mean_depth],2), " m"), size = 2)


}

ggsave("arranged_areaplots_classified.pdf", marrangeGrob(grobs = lme_areaplots_classified, ncol=2, nrow = 3))
```


PCA

- size
- region (color)
- mean depth
- mid-latitude
- classification (color)
- max depth

```{r PCA}
library(ggfortify)
#get rid of antartica and arctic (can remove this next round)
LME_bathy_statistics <- LME_bathy_statistics[lme_number != 61,][lme_number != 64,]


LME.pca <- prcomp(LME_bathy_statistics[,.(lme_area, mean_depth, max_depth, center_latitude)], scale. = T)


LME.plot_type <- autoplot(LME.pca, data = LME_bathy_statistics, colour = 'type',
                          loadings = TRUE, loadings.colour = 'black',
                          loadings.label = TRUE, loadings.label.size = 4) + theme_classic()
LME.plot_region <- autoplot(LME.pca, data = LME_bathy_statistics, colour = 'region',
                          loadings = TRUE, loadings.colour = 'black',
                          loadings.label = TRUE, loadings.label.size = 4) + theme_classic()

LME.plot_type
LME.plot_region

ggsave(plot = LME.plot_type, filename = "LME.plot_type.pdf")
ggsave(plot = LME.plot_region, filename = "LME.plot_region.pdf")

```
How many of each type
Multimodal                   Mid Dominant            Shallow Dominant 
       44 (69%)                8 (13%)                12 (20%) 
```{r how many of each type}
table(LME_bathy_statistics$type)
44/64 #multimodal
8/64 #mid-dominant
12/64 #shallow-dominant
```
