---
title: "Starting with Bathy for Shelves"
output: html_notebook
---

```{r setup}
library(raster)
library(rgdal)
library(sf)
library(ncdf4)
library(rmapshaper)
library(tidyverse)
library(diptest)
library(moments)
library(viridis) #colors
library(data.table)
library(hydroTSM) #hypsometric curves
library(gridExtra)
library(ggfortify)

etopo_shelf_df <- readRDS("~/Documents/grad school/Rutgers/Repositories/shelf_habitat_distribution/etopo_shelf_df.rds")
#bring in bathy for shelf regions

#LMEs
LME_spdf <- readOGR("LME66/LMEs66.shp")
head(LME_spdf)
```

Let's just start with one LME to get started
```{r LME_baltic}
LME_baltic <- LME_spdf[1,]
plot(LME_baltic)


```


Make bathy layer into raster
```{r rasterize}
proj <- crs(LME_spdf)
etopo_shelf_raster <- rasterFromXYZ(etopo_shelf_df, crs = proj)

#make figure of shelves
  #png("etopo_shelf_bathy.png", height=5, width=7, units="in", res=300)
  #plot(etopo_shelf_raster)
  #dev.off()

```

Clip raster to each individual polygon

```{r clip raster to polygon}
#crop bathy layer to LME subset
bathy_baltic_extent <- crop(etopo_shelf_raster, extent(LME_baltic))

#which areas of raster fall within borders?
bathy_baltic_mask <- mask(bathy_baltic_extent, LME_baltic)

#convert to df for plotting

  # First, to a SpatialPointsDataFrame
  bathy_baltic_pts <- rasterToPoints(bathy_baltic_mask, spatial = TRUE)
  # Then to a 'conventional' dataframe
  bathy_baltic_dt  <- data.table(data.frame(bathy_baltic_pts))

#keep depth positive  
bathy_baltic_dt[,depth := -(layer)]

#fortify to plot shapefile on top
#LME_baltic_df <- fortify(LME_baltic) (don't need because indstead I make polygon to surround raster)

#To get a polygon that surrounds cells that are not NA

  # make all values the same.
  r <- reclassify(bathy_baltic_mask, cbind(-Inf, Inf, 1))
  
  # convert to polygons (you need to have package 'rgeos' installed for this to work)
  pp <- rasterToPolygons(r, dissolve=TRUE)
  
  LME_reduced_baltic_df <- fortify(pp)

ggplot(bathy_baltic_dt) +
  geom_tile(aes(x = x, y = y, fill = depth)) +
  scale_fill_viridis(direction = -1, limits = c(0,2000)) +
  theme_classic() +
  labs(x = "Longitude", y = "Latitude", fill = "Depth (m)") +
  guides(fill = guide_colorbar(reverse = T)) +
  geom_path(data = LME_reduced_baltic_df, aes(x = long, y = lat, group = group), fill = NA, color = "black")

summary(bathy_baltic_dt)

#how to get area of raster

  #get sizes of all cells in raster [km2]
  cell_size<-area(bathy_baltic_mask, na.rm=TRUE, weights=FALSE)
  #delete NAs from vector of all raster cells
  ##NAs lie outside of the rastered region, can thus be omitted
  cell_size<-cell_size[!is.na(cell_size)]
  #compute area [km2] of all cells in geo_raster
  bathy_raster_area<-length(cell_size)*median(cell_size)

```
Will have to be creative on how to map LMEs that cross 180 longitude, ex. Aleutian Islands
```{r high longitude examples}
LME_datatable <- LME_spdf@data
LME_aleutian <- LME_spdf[59,]

#crop bathy layer to LME subset
bathy_aleutian_extent <- crop(etopo_shelf_raster, extent(LME_aleutian))

#which areas of raster fall within borders?
bathy_aleutian_mask <- mask(bathy_aleutian_extent, LME_aleutian)

#convert to df for plotting

  # First, to a SpatialPointsDataFrame
  bathy_aleutian_pts <- rasterToPoints(bathy_aleutian_mask, spatial = TRUE)
  # Then to a 'conventional' dataframe
  bathy_aleutian_dt  <- data.table(data.frame(bathy_aleutian_pts))

#keep depth positive  
bathy_aleutian_dt[,depth := -(layer)]

bathy_aleutian_dt$x_shift <- ifelse(bathy_aleutian_dt$x >0, bathy_aleutian_dt$x - 360, bathy_aleutian_dt$x)


#To get a polygon that surrounds cells that are not NA

  # make all values the same.
  r <- reclassify(bathy_aleutian_mask, cbind(-Inf, Inf, 1))
  
  # convert to polygons (you need to have package 'rgeos' installed for this to work)
  pp <- rasterToPolygons(r, dissolve=TRUE)
  
  LME_reduced_aleutian_df <- fortify(pp)
  
  
LME_reduced_aleutian_df$long_shift <- ifelse(LME_reduced_aleutian_df$long >0, LME_reduced_aleutian_df$long - 360, LME_reduced_aleutian_df$long)
  

ggplot(bathy_aleutian_dt) +
  geom_tile(aes(x = x_shift, y = y, fill = depth)) +
  scale_fill_viridis(direction = -1, limits = c(0,2000)) +
  theme_classic() +
  labs(x = "Longitude (˚E)", y = "Latitude", fill = "Depth (m)") +
  guides(fill = guide_colorbar(reverse = T)) +
  geom_path(data = LME_reduced_aleutian_df, aes(x = long_shift, y = lat, group = group), fill = NA, color = "black")

```

Hypsometric Curves or Density Curve?
```{r hyposometric curves}
#list of values within raster
bathy_baltic_depth_list <- getValues(bathy_baltic_mask)
bathy_baltic_depth_list <- bathy_baltic_depth_list[!is.na(bathy_baltic_depth_list)]

bathy_baltic_depth_list_pos <- -1*bathy_baltic_depth_list

bathy_baltic_SGDF <- as(bathy_baltic_mask, 'SpatialGridDataFrame') # as spatial grid data frame
hypsometric(bathy_baltic_SGDF)

#dimensions of raster?
dim(bathy_baltic_mask)

#total number of cells?
cells <- dim(bathy_baltic_mask)[1]*dim(bathy_baltic_mask)[2]


#cells of each depth value in data table
bathy_baltic_bydepth <- data.table(freq(bathy_baltic_mask))

#NA cells
na_cell_count <- bathy_baltic_bydepth[is.na(value),]$count

#filled cells

cells_filled <- cells-na_cell_count

bathy_baltic_bydepth_complete <- bathy_baltic_bydepth[!is.na(value)]

#calculate percent of total area in polygon
bathy_baltic_bydepth_complete[,percent_area := count/sum(count)]


bathy_baltic_bydepth_complete[,area := percent_area*bathy_raster_area][,depth := -value]



(plot <- ggplot(data = bathy_baltic_bydepth_complete, aes(x = depth, y = area, fill = depth)) +
  geom_col(width = 1) +
  theme_classic() +
  labs(x = "Depth (m)", y = "Area (km^2)", fill = "Depth (m)") +
  scale_y_continuous(expand = c(0, 0)) +
  scale_x_continuous(expand = c(0,0)) +
  scale_fill_gradientn(colors = rev(viridis(10)), limits = c(0, 2000)) +
  guides(position = "bottom", fill = guide_colorbar(reverse = T)))

#alternative, simple density plot
ggplot(data = bathy_baltic_mask, aes(x = -layer, fill = -layer)) +
  geom_density() +
  theme_classic()
  

ggsave(plot, filename = "Baltic_Sea.png")
```
What if we model instead of use true data? Instead, I need a density curve!!!

```{r model versus data}
ggplot(data = bathy_baltic_bydepth_complete, aes(x = depth, y = count, fill = depth)) +
  #geom_col(width = 1) +
  geom_density() +
  geom_smooth(method = "gam", se = F) +
  theme_classic() +
  labs(x = "Depth (m)", y = "Area (km^2)", fill = "Depth (m)") +
  scale_y_continuous(expand = c(0, 0)) +
  scale_x_continuous(expand = c(0,0)) +
  scale_fill_gradientn(colors = rev(viridis(5)), limits = c(0, 2000)) +
  guides(position = "bottom", fill = guide_colorbar(reverse = T))

library(mgcv)
gam_model <- gam(area ~ s(depth, bs = "cr"), data = bathy_baltic_bydepth_complete)

#cr = cubic spline 

predicted_values <- predict(gam_model)

bathy_baltic_bydepth_complete[, preds := predicted_values]

ggplot(data = bathy_baltic_bydepth_complete, aes(x = depth, y = area, fill = depth)) +
  geom_col(width = 1) +
  geom_density() +
  #geom_smooth(method = "gam", se = F) +
  geom_line(aes(x = depth, y = preds)) +
  theme_classic() +
  labs(x = "Depth (m)", y = "Area (km^2)", fill = "Depth (m)") +
  scale_y_continuous(expand = c(0, 0)) +
  scale_x_continuous(expand = c(0,0)) +
  scale_fill_gradientn(colors = rev(viridis(10)), limits = c(0, 2000)) +
  guides(position = "bottom", fill = guide_colorbar(reverse = T))


```


How to classify different LMES? Dip (Modality) & Skew
 "distributions with a dip value >0.01 and with significant (p<0.05) deviations from unimodality to the hourglass classification. 
 
 For distributions with a dip value ≤0.01, we assigned those with a Type-I skewness ≥0.5 to pyramid, 
 
 those with skewness ≤ −0.5 to inverse pyramid, 
 
 and the remainder to diamond, representing those with approximately normal distributions. 
 
 We chose skew cutoffs of 0.5 and −0.5 to capture right- and left-skewed distributions, respectively, and to bound distributions approximating symmetry."
 
Modality: In mathematics, unimodality means possessing a unique mode. More generally, unimodality means there is only a single highest value, somehow defined, of some mathematical object. (Wikipedia)

Skew: In probability theory and statistics, skewness is a measure of the asymmetry of the probability distribution of a real-valued random variable about its mean. The skewness value can be positive, zero, negative, or undefined. (Wikipedia)
 
```{r dip test for unimodality}
#Computes Hartigans’ dip test statistic for testing unimodality, and additionally the modal interval.

diptest <- dip.test(bathy_baltic_depth_list_pos, simulate.p.value = TRUE, B = 2000)
p.value <- diptest$p.value

```

Skew

* If skewness is less than -1 or greater than 1, the distribution is highly skewed.
* If skewness is between -1 and -0.5 or between 0.5 and 1, the distribution is moderately skewed.
* If skewness is between -0.5 and 0.5, the distribution is approximately symmetric.
```{r skew}
skew <- skewness(bathy_baltic_depth_list_pos, na.rm = T)
```


Maybe kurtosis too?

Kurtosis is a statistical measure that defines how heavily the tails of a distribution differ from the tails of a normal distribution. In other words, kurtosis identifies whether the tails of a given distribution contain extreme values.
```{r kurtosis}
kurtosis <- kurtosis(bathy_baltic_depth_list_pos, na.rm = T)
```

Mean depth value
```{r mean depth}
mean <- mean(bathy_baltic_depth_list_pos)
```

Maybe then do a MDS including kurtosis, skew, and modality measures for every LME

See how many groups it goes into?

Shallow dominant: High positive skew, unimodal
Mid-Dominant: Relatively normal (low skew, kurtosis = 0, unimodal)
Deep-Dominant: High negative skew, unimodal
Uniform: kurtosis < -1.2, unimodal, little skew
Hourglass: Bimodality

Text to add to plot:
 - class
 - skew
 - dip
 - mean
 - bottleneck

```{r full plot}

plot +
  annotate("text", x = 300, y = 4000, label = paste0("Dip Test: ", signif(diptest$statistic,2), " P-value = ", signif(p.value))) +
  annotate("text", x = 300, y = 3700, label = paste0("Skew: ", round(skew,2))) +
  annotate("text", x = 300, y = 3400, label = paste0("Kurtosis: ",round(kurtosis,2))) +
  annotate("text", x = 300, y = 3100, label = paste0("Mean: ",round(mean,2), " m")) +
  geom_vline(xintercept = mean, color = "red")
```

--------
Now let's make a loop!!

```{r loop}

#plot map, plot hypsograph, plot histogram

#reorder LME_spdf by LME # 
LME_spdf_ordered <- LME_spdf[order(LME_spdf$LME_NUMBER),]


LME_bathy_statistics <- as.data.table(matrix(nrow = nrow(LME_spdf)))
LME_bathy_statistics[, lme_name := as.factor(V1)][, lme_number := as.numeric(V1)][,lme_area := as.numeric(V1)][, dip_test := as.numeric(V1)][, dip_p_value := as.numeric(V1)][, skew := as.numeric(V1)][, kurtosis := as.numeric(V1)][, mean_depth := as.numeric(V1)]
LME_bathy_statistics[, V1 := NULL]


lme_maps <- vector("list", length = nrow(LME_spdf))
lme_densityplots <- vector("list", length = nrow(LME_spdf))
lme_areaplots <- vector("list", length = nrow(LME_spdf))

for (i in 1:nrow(LME_bathy_statistics)) {
  LME_single <- LME_spdf_ordered[i,]
  LME_bathy_statistics[i, "lme_name"] <- LME_single$LME_NAME
  LME_bathy_statistics[i, "lme_number"] <- LME_single$LME_NUMBER
  
  #clip raster to LME
  LME_extent <- crop(etopo_shelf_raster, extent(LME_single))
  
  #which areas of raster fall within borders?
  LME_mask <- mask(LME_extent, LME_single)
  
  #conver to df for plotting
    # First, to a SpatialPointsDataFrame
    LME_points <- rasterToPoints(LME_mask, spatial = TRUE)
    # Then to a 'conventional' dataframe
    LME_dt  <- data.table(data.frame(LME_points))
  
  #keep depth positive  
  LME_dt[,depth := -(layer)]
  
  LME_dt$x_shift <- ifelse(
    max(LME_dt$x)-min(LME_dt$x) > 200 & LME_dt$x >0,
    LME_dt$x - 360,
    LME_dt$x)
  
  #size of raster  
  #get sizes of all cells in raster [km2]
    cell_size<-area(LME_mask, na.rm=TRUE, weights=FALSE)
    #delete NAs from vector of all raster cells
    ##NAs lie outside of the rastered region, can thus be omitted
    cell_size<-cell_size[!is.na(cell_size)]
    #compute area [km2] of all cells in geo_raster
    bathy_raster_area<-length(cell_size)*median(cell_size)
    
    LME_bathy_statistics[i, "lme_area"] <- bathy_raster_area
    
  #To get a polygon that surrounds cells that are not NA
  
    # make all values the same.
    r <- reclassify(LME_mask, cbind(-Inf, Inf, 1))
    
    # convert to polygons (you need to have package 'rgeos' installed for this to work)
    pp <- rasterToPolygons(r, dissolve=TRUE)
    
    LME_reduced_df <- fortify(pp)
    
    LME_reduced_df$long_shift <- ifelse(
    max(LME_reduced_df$long)-min(LME_reduced_df$long) > 200 & LME_reduced_df$long >0,
    LME_reduced_df$long - 360,
    LME_reduced_df$long)
  
  lme_maps[[i]] <- ggplot(LME_dt) +
    geom_tile(aes(x = x_shift, y = y, fill = depth)) +
    scale_fill_viridis(direction = -1, limits = c(0,2000)) +
    theme_classic() +
    labs(x = "Longitude", y = "Latitude", fill = "Depth (m)") +
    guides(fill = guide_colorbar(reverse = T)) +
    geom_path(data = LME_reduced_df, aes(x = long_shift, y = lat, group = group), color = "black") +
    annotate("text", x=Inf, y = max(LME_reduced_df$lat)+1, hjust=1, label = paste0("LME ", LME_bathy_statistics[i,lme_number], ": ", LME_bathy_statistics[i,lme_name]), fontface = "bold", size = 2.5)
  
  
    #list of values within raster
  LME_bathy_depth_list <- getValues(LME_mask)
  LME_bathy_depth_list <- LME_bathy_depth_list[!is.na(LME_bathy_depth_list)]
  
  LME_bathy_depth_list_pos <- -1*LME_bathy_depth_list


  #cells of each depth value in data table
  LME_bathy_bydepth <- data.table(freq(LME_mask))
  
  LME_bathy_bydepth_complete <- LME_bathy_bydepth[!is.na(value)]
  
  #calculate percent of total area in polygon
  LME_bathy_bydepth_complete[,percent_area := count/sum(count)]
  
  
  LME_bathy_bydepth_complete[,area := percent_area*bathy_raster_area][,depth := -value]
  
  
  
  plot <- ggplot(data = LME_bathy_bydepth_complete, aes(x = depth, y = area, fill = depth)) +
    geom_col(width = 1) +
    theme_classic() +
    labs(x = "Depth (m)", y = "Area (km^2)", fill = "Depth (m)") +
    scale_y_continuous(expand = c(0, 0)) +
    scale_x_continuous(expand = c(0,0)) +
    scale_fill_gradientn(colors = rev(viridis(5)), limits = c(0, 2000)) +
    guides(position = "bottom", fill = guide_colorbar(reverse = T))
  
  
  #unimodality
  #Computes Hartigans’ dip test statistic for testing unimodality, and additionally the modal interval.
  
  diptest <- dip.test(LME_bathy_depth_list_pos, simulate.p.value = TRUE, B = 2000)
  LME_bathy_statistics[i,"dip_test"] <- diptest$statistic[1]
  p.value <- diptest$p.value
  LME_bathy_statistics[i,"dip_p_value"] <- p.value
  skew <- skewness(LME_bathy_depth_list_pos, na.rm = T)
  LME_bathy_statistics[i,"skew"] <- skew
  kurtosis <- kurtosis(LME_bathy_depth_list_pos, na.rm = T)
  LME_bathy_statistics[i,"kurtosis"] <- kurtosis
  mean <- mean(LME_bathy_depth_list_pos)
  LME_bathy_statistics[i,"mean_depth"] <- mean
  
  max_area <- max(LME_bathy_bydepth_complete$area)
  
  #full plot
  lme_areaplots[[i]] <- plot +
    geom_vline(xintercept = mean, color = "red") +
    annotate("text", x=Inf, y = max_area-max_area/20, hjust=1, label = paste0(LME_bathy_statistics[i,lme_name]), fontface = "bold", size = 2) +
    annotate("text", x=Inf, y = max_area-2*(max_area/20), hjust=1, label = paste0("Dip Test: ", signif(diptest$statistic,2), " P-value = ", signif(p.value, 2)), size = 2) +
    annotate("text", x=Inf, y = max_area-3*(max_area/20), hjust=1, label = paste0("Skew: ", round(skew,2)), size = 2) +
    annotate("text", x=Inf, y = max_area-4*(max_area/20), hjust=1, label = paste0("Kurtosis: ",round(kurtosis,2)), size = 2) +
    annotate("text", x=Inf, y = max_area-5*(max_area/20), hjust=1, label = paste0("Mean: ",round(mean,2), " m"), size = 2)
  
  #density values, find max to help with annotation
  density_distribution <- density(LME_bathy_depth_list_pos)
  max_density <- max(density_distribution$y)
  
lme_densityplots[[i]] <- ggplot() +
    geom_density(aes(LME_bathy_depth_list_pos)) +
    annotate("text", x=Inf, y = max_density+0.003, hjust=1, label = paste0(LME_bathy_statistics[i,lme_name]), fontface = "bold", size = 2) +
    theme_classic() +
    labs(x = "Depth (m)", y = "Density")
  
    print(paste(i, nrow(LME_bathy_statistics), sep = "/"))
  

}

saveRDS(LME_bathy_statistics, file = "LME_bathy_statistics.RData")
saveRDS(lme_densityplots, file = "lme_densityplots.Rdata")
saveRDS(lme_areaplots, file = "lme_areaplots.Rdata")
saveRDS(lme_maps, file = "lme_maps.Rdata")

#plots to pdf
ggsave("arranged_maps.pdf", marrangeGrob(lme_maps, ncol=2, nrow = 2))
ggsave("arranged_densityplots.pdf", marrangeGrob(grobs = lme_densityplots, ncol=2, nrow = 3))
ggsave("arranged_areaplots.pdf", marrangeGrob(grobs = lme_areaplots, ncol=2, nrow = 3))
```

PCA
```{r PCA}

LME.pca <- prcomp(LME_bathy_statistics[,c(4,6,7)])
LME.plot <- autoplot(LME.pca)
```

TO DO:
make plots into curves instead of actual data** GAM, or kernel density? maybe neither

Add classifications to data frame

* If skewness is less than -1 or greater than 1, the distribution is highly skewed.
* If skewness is between -1 and -0.5 or between 0.5 and 1, the distribution is moderately skewed.
* If skewness is between -0.5 and 0.5, the distribution is approximately symmetric.

distributions with a dip value >0.01 and with significant (p<0.05) deviations from unimodality to the hourglass classification. 

For distributions with a dip value ≤0.01, we assigned those with a Type-I skewness ≥0.5 to shallow dominant, those with skewness ≤ −0.5 to deep dominant, and the remainder to diamond, representing those with approximately normal distributions. 

```{r expanded dataframe}
LME_bathy_statistics[, skew_descriptor := ifelse(abs(skew) > 1, "high_skew", ifelse(abs(skew) < 0.5, "symmetric", "mod_skew"))]

LME_bathy_statistics[, hourglass := ifelse(dip_test > 0.01, ifelse(p.value < 0.05, T, F), F)]

LME_bathy_statistics[, deep_dominant := ifelse(dip_test <= 0.01, 
                                               ifelse(skew <= -0.5, T, F), F)]

LME_bathy_statistics[, shallow_dominant := ifelse(dip_test <= 0.01, 
                                                  ifelse(skew >= 0.5, T, F), F)]

LME_bathy_statistics[, mid_dominant := ifelse(abs(skew) <0.5, T, F)]

#negative kurtosis (no tails)
LME_bathy_statistics[, uniform := ifelse(kurtosis < 0, T, F)]

```

Are all regions labeled by type?
```{r labeling check}
LME_bathy_statistics[,type := ifelse(hourglass == T, "hourglass",
                                    ifelse(deep_dominant == T, "deep_dominant",
                                           ifelse(shallow_dominant == T, "shallow_dominant",
                                                  ifelse(mid_dominant == T, "mid_dominant", 
                                                         ifelse(uniform == T, "uniform", "none")))))]
```

PCA
- size
- region



BIG Qs to ASK MALIN
 - do I need to limit LMEs to continental shelf regions? does blue habitat seem like a good limiter?
 Becca thinks Blue Habitat looks good, so good enough to me
 - https://www.bluehabitats.org/
 - should I instead go with 200 m like Hailey did? Instead, I am going to use Blue Habitat, and limit to 2000m depth