---
title: "Starting with Bathy for Shelves"
output: html_notebook
---

```{r setup}
library(raster)
library(rgdal)
library(sf)
library(ncdf4)
library(rmapshaper)
library(tidyverse)
library(diptest)
library(moments)
library(viridis) #colors
library(data.table)
library(hydroTSM) #hypsometric curves

etopo_shelf_df <- readRDS("~/Documents/grad school/Rutgers/Repositories/shelf_habitat_distribution/etopo_shelf_df.rds")
#bring in bathy for shelf regions

#LMEs
LME_spdf <- readOGR("LME66/LMEs66.shp")
head(LME_spdf)
```

Let's just start with one LME to get started
```{r LME_baltic}
LME_baltic <- LME_spdf[1,]
plot(LME_baltic)

#rather than a spatial polygon, we need points within that polygon
LME_baltic_points <- spsample(LME_baltic, n = 30000, type = "regular")
LME_baltic_points_xy <- geom(LME_baltic_points)[,2:3]

```


Make bathy layer into raster
```{r rasterize}
proj <- crs(LME_spdf)
etopo_shelf_raster <- rasterFromXYZ(etopo_shelf_df, crs = proj)

#make figure of shelves
png("etopo_shelf_bathy.png", height=5, width=7, units="in", res=300)
plot(etopo_shelf_raster)
dev.off()

```

Clip raster to each individual polygon

```{r clip raster to polygon}
#crop bathy layer to LME subset
bathy_baltic_extent <- crop(etopo_shelf_raster, extent(LME_baltic))

#which areas of raster fall within borders?
bathy_baltic_mask <- mask(bathy_baltic_extent, LME_baltic)

#convert to df for plotting

  # First, to a SpatialPointsDataFrame
  bathy_baltic_pts <- rasterToPoints(bathy_baltic_mask, spatial = TRUE)
  # Then to a 'conventional' dataframe
  bathy_baltic_dt  <- data.table(data.frame(bathy_baltic_pts))

#keep depth positive  
bathy_baltic_dt[,depth := -(layer)]

ggplot(bathy_baltic_dt, aes(x = x, y = y, fill = depth)) +
  geom_tile() +
  scale_fill_viridis(direction = -1) +
  theme_classic() +
  labs(x = "Longitude", y = "Latitude", fill = "Depth (m)") +
  guides(fill = guide_colorbar(reverse = T))
```

Hypsometric Curves
```{r hyposometric curves}

bathy_baltic_SGDF <- as(bathy_baltic_mask, 'SpatialGridDataFrame') # as spatial grid data frame
hypsometric(bathy_baltic_SGDF)

(plot <- ggplot(data = bathy_baltic_dt[!is.na(depth),], aes(x = depth, fill=..x..)) +
  geom_histogram(binwidth = 5) +
  theme_classic() +
  labs(x = "Depth (m)", y = "Frequency") +
  scale_fill_gradientn(colors = rev(viridis(5))) +
  guides(fill = guide_colorbar(reverse = T)))

ggsave(plot, filename = "Baltic_Sea.png")
```


How to classify different LMES? Dip (Modality) & Skew
 "distributions with a dip value >0.01 and with significant (p<0.05) deviations from unimodality to the hourglass classification. 
 
 For distributions with a dip value ≤0.01, we assigned those with a Type-I skewness ≥0.5 to pyramid, 
 
 those with skewness ≤ −0.5 to inverse pyramid, 
 
 and the remainder to diamond, representing those with approximately normal distributions. 
 
 We chose skew cutoffs of 0.5 and −0.5 to capture right- and left-skewed distributions, respectively, and to bound distributions approximating symmetry."
 
Modality: In mathematics, unimodality means possessing a unique mode. More generally, unimodality means there is only a single highest value, somehow defined, of some mathematical object. (Wikipedia)

Skew: In probability theory and statistics, skewness is a measure of the asymmetry of the probability distribution of a real-valued random variable about its mean. The skewness value can be positive, zero, negative, or undefined. (Wikipedia)
 
```{r dip test for unimodality}
#Computes Hartigans’ dip test statistic for testing unimodality, and additionally the modal interval.

diptest <- dip.test(bathy_baltic_dt$depth, simulate.p.value = FALSE, B = 2000)


```

Skew

* If skewness is less than -1 or greater than 1, the distribution is highly skewed.
* If skewness is between -1 and -0.5 or between 0.5 and 1, the distribution is moderately skewed.
* If skewness is between -0.5 and 0.5, the distribution is approximately symmetric.
```{r skew}
skew <- skewness(bathy_baltic_dt$depth, na.rm = T)
```


Maybe kurtosis too?

Kurtosis is a statistical measure that defines how heavily the tails of a distribution differ from the tails of a normal distribution. In other words, kurtosis identifies whether the tails of a given distribution contain extreme values.
```{r kurtosis}
kurtosis <- kurtosis(bathy_baltic_dt$depth, na.rm = T)
```

Maybe then do a MDS including kurtosis, skew, and modality measures for every LME

See how many groups it goes into?

Shallow dominant: High positive skew, unimodal
Mid-Dominant: Relatively normal (low skew, kurtosis = 0, unimodal)
Deep-Dominant: High negative skew, unimodal
Uniform: kurtosis < -1.2, unimodal, little skew
Hourglass: Bimodality


--------
Now let's make a loop!!