---
title: "2˚ Shifts global area change analysis"
output: html_notebook
---

```{r setup}
library(raster)
library(sf)
library(ncdf4)
library(rmapshaper)
library(tidyverse)
library(diptest)
library(moments)
library(viridis) #colors
library(data.table)
library(hydroTSM) #hypsometric curves
library(gridExtra)
library(maptools)
library(rgdal)
library(rgeos)
library(SpaDES)
library(rnaturalearth)
library(rnaturalearthdata)

load(here::here("output","2_degree_latitude","west_pac_shelf_areas_2degrees.Rdata"))
load(here::here("output","2_degree_latitude","west_atl_shelf_areas_2degrees.Rdata"))
load(here::here("output","2_degree_latitude","west_ind_shelf_areas_2degrees.Rdata"))
load(here::here("output","2_degree_latitude","east_pac_shelf_areas_2degrees.Rdata"))
load(here::here("output","2_degree_latitude","east_atl_shelf_areas_2degrees.Rdata"))
east_ind_shelf_areas <- readRDS(here::here("output","2_degree_latitude","east_ind_shelf_areas_2degrees.rds"))

```

For ease, merge areas (I need to do this when I actually have matched them in degree_shifts_projected_2degrees script)
```{r merge tables for each region}
west_pac_shelf_areas.r <- west_pac_shelf_areas[,1:3][,region := "west_pac"]
west_atl_shelf_areas.r <- west_atl_shelf_areas[,1:3][,region := "west_atl"]
west_ind_shelf_areas.r <- west_ind_shelf_areas[,1:3][,region := "west_ind"]
east_pac_shelf_areas.r <- east_pac_shelf_areas[,1:3][,region := "east_pac"]
east_atl_shelf_areas.r <- east_atl_shelf_areas[,1:3][,region := "east_atl"]
east_ind_shelf_areas.r <- east_ind_shelf_areas[,1:3][,region := "east_ind"]



shelf_areas <- rbind(west_pac_shelf_areas.r, west_atl_shelf_areas.r, west_ind_shelf_areas.r, east_pac_shelf_areas.r, east_atl_shelf_areas.r, east_ind_shelf_areas.r)

shelf_areas[,hemisphere := ifelse(latitude_end >0, "Northern", "Southern")]

#be sure to order table first so that shift is calculated accurately

#include absolute value of latitude end to help with sorting
shelf_areas[,latitude_end_abs  := abs(latitude_end)]
setkey(shelf_areas, region, hemisphere, latitude_end_abs) #sort by region, and then hemisphere, and then latitude_end_abs increasing (because we want to be moving outward from the equator)

#include mid_latitude value for plotting
shelf_areas[,latitude_mid  := (latitude_start+latitude_end)/2]



```


Calculate change in area from one bin to the next
```{r change in area}
#area change
shelf_areas[, area_change :=  area_rasterarea -  data.table::shift(area_rasterarea, n = 1, type = "lag"), .(region, hemisphere)] #change in species from previous bin to this bin?
#note that it's important to include grouping for region so that calculations are correct, and there's always a gain at the lowest latitude

#positive versus negative change

shelf_areas[,change_dir := ifelse(area_change > 0, "positive", "negative")]

#add labels to regions
shelf_areas[,region := factor(region, levels = c("east_atl", "east_ind", "east_pac", "west_atl", "west_ind", "west_pac"), labels = c("East Atlantic","East Indian","East Pacific","West Atlantic","West Indian","West Pacific"))]

(area_change_latitude  <- ggplot(data = shelf_areas) +
  geom_col(aes(x=latitude_end, y=area_change/1000, fill = change_dir)) +
 labs(x=paste0("Latitude ","\u00B0","E"), y = expression(paste("Area (1000s of ", km^{2},")"))) +
     scale_fill_viridis_d(option = "A", begin = 0.3, end = 0.7, name = "Direction of Shelf\nArea Change\nWith 2\u00B0 Shift", labels = c("Negative", "Positive")) +
  geom_vline(xintercept = 0) +
  geom_hline(yintercept = 0) +
  coord_flip() +
  theme_classic() +
   facet_wrap(~region, scales = "free"))
  

ggsave(area_change_latitude, file = "area_change_latitude.jpg")

```

Let's make this same figure but for species lost or gained. 

Make datatable with statistics
```{r datatable with stats}
shelf_areas[, sum_change := sum(area_change, na.rm = T)] #total change in area across all shelf regions
shelf_areas[, sum_change_reg := sum(area_change, na.rm = T), region] #total change in area by region
shelf_areas[, sum_change_hemis := sum(area_change, na.rm = T), hemisphere] #total change in area by hemisphere
shelf_areas[, sum_change_hemis_reg := sum(area_change, na.rm = T), .(region, hemisphere)] #total change in area by both hemisphere and region

shelf_areas[, original_area_reg := sum(area_rasterarea, na.rm = T), region] #original area by region
shelf_areas[, final_area_reg := original_area_reg + sum_change_reg, region] #final area by region

(shelf_area_change_all_stats <- unique(shelf_areas[,.(region, hemisphere, sum_change, sum_change_reg, sum_change_hemis, sum_change_hemis_reg)]))

(shelf_area_change_region_stats <- unique(shelf_areas[,.(region, sum_change_reg)]))
(shelf_area_change_hemis_stats <- unique(shelf_areas[,.(hemisphere, sum_change_hemis)]))
(shelf_area_change_hemis_reg_stats <- unique(shelf_areas[,.(hemisphere, region, sum_change_hemis_reg)]))

#color by direction
shelf_area_change_hemis_reg_stats[,change_direction := factor(ifelse(sum_change_hemis_reg >1, "Gain", "Loss"))]

#individual bins
shelf_areas[, original_spp_predicted_meixler :=
              exp(6.854 + (0.185 * log(area_rasterarea)) -  (1.084 * log(abs(0.5*(latitude_start+latitude_end)))))]

shelf_areas[, original_spp_predicted_levin := 16.18*area_rasterarea^0.226]

#be sure to order table so that shift is used accurately (I did this above also, this is just a check)
#include absolute value of latitude end
setkey(shelf_areas, region, hemisphere, latitude_end_abs)

#species change
shelf_areas[, change_spp_predicted_meixler :=  original_spp_predicted_meixler -  data.table::shift(original_spp_predicted_meixler, n = 1, type = "lag"), .(region, hemisphere)] #change in species from previous bin to this bin?
#note that it's important to include grouping for region so that calculations are correct, and there's always a gain at the lowest latitude

shelf_areas[, change_spp_predicted_levin :=  original_spp_predicted_levin -  data.table::shift(original_spp_predicted_levin, n = 1, type = "lag"), .(region, hemisphere)] #change in species from previous bin to this bin?

#percent change
shelf_areas[, percent_change_richness := (original_spp_predicted_levin-data.table::shift(original_spp_predicted_levin, type = "lag"))/data.table::shift(original_spp_predicted_levin, type = "lag"), .(region,hemisphere)]
shelf_areas[,spp_per_change_significant := ifelse(percent_change_richness >= 1, "Gain",ifelse(percent_change_richness <= -0.5, "Loss","NA"))]

#new data table with only these values
shelf_areas_highlight <- shelf_areas[spp_per_change_significant == "Gain" | spp_per_change_significant == "Loss"]

#color code by direction of spp gain or loss
shelf_areas[, spp_change_pos_neg_meixler := ifelse(change_spp_predicted_meixler > 0, "Positive","Negative")]
shelf_areas[, spp_change_pos_neg_levin := ifelse(change_spp_predicted_levin > 0, "Positive","Negative")]

#full stats for spp change
shelf_areas[,spp_change_levin_by_reg := sum(change_spp_predicted_levin, na.rm = T), region]
shelf_areas[,spp_change_levin_by_hemis := sum(change_spp_predicted_levin, na.rm = T), hemisphere]
shelf_areas[,spp_change_levin_by_reg_hemis := sum(change_spp_predicted_levin, na.rm = T), .(region, hemisphere)]

#summary for spp change
(shelf_spp_change_region_stats <- unique(shelf_areas[,.(region, spp_change_levin_by_reg)]))
(shelf_spp_change_hemis_stats <- unique(shelf_areas[,.(hemisphere, spp_change_levin_by_hemis)]))
(shelf_spp_change_hemis_reg_stats <- unique(shelf_areas[,.(hemisphere, region, spp_change_levin_by_reg_hemis)]))

#color by direction
shelf_spp_change_hemis_reg_stats[,change_direction := factor(ifelse(spp_change_levin_by_reg_hemis > 0, "Gain", "Loss"))] 

#save data table
saveRDS(shelf_areas, file = here::here("output","2_degree_latitude","shelf_areas.rds"))
  
```
Figure for spp richness per latitude
```{r spp richness per latitude}
(spp_change_latitude  <- ggplot(data = shelf_areas, aes(x=latitude_mid, y=original_spp_predicted_levin)) +
   geom_area(fill = "azure3") +
 labs(x=paste0("Latitude ","\u00B0","E"), y = "Species Richness") +
  geom_vline(xintercept = 0) +
  geom_hline(yintercept = 0) +
  coord_flip() +
  theme_classic() +
   facet_wrap(~region, scales = "free"))
```
Mark 50% loss or 200% gain of species (this isn't useful)
```{r gain loss of substantial # of species}
(spp_change_latitude  <- ggplot() +
   geom_area(data = shelf_areas, aes(x=latitude_mid, y = original_spp_predicted_levin), fill = "azure3") +
   geom_rug(data = shelf_areas_highlight, aes(x=latitude_mid, color = spp_per_change_significant)) +
 labs(x=paste0("Latitude ","\u00B0","E"), y = "Species Richness") +
  geom_vline(xintercept = 0) +
  geom_hline(yintercept = 0) +
  coord_flip() +
  theme_classic() +
   facet_wrap(~region, scales = "free"))
```


Figure for spp richness change by latitude
```{r spp change lat}
(spp_change_latitude  <- ggplot(data = shelf_areas) +
  geom_col(aes(x=latitude_end, y=change_spp_predicted_levin, fill = change_dir)) +
 labs(x=paste0("Latitude ","\u00B0","E"), y = "Change in Richness\nwith 2\u00B0 Shift") +
     scale_fill_viridis_d(option = "A", begin = 0.3, end = 0.7, name = "Direction of\nRichness Change\nWith 2\u00B0 Shift", labels = c("Negative", "Positive")) +
  geom_vline(xintercept = 0) +
  geom_hline(yintercept = 0) +
  coord_flip() +
  theme_classic() +
   facet_wrap(~region, scales = "free"))
```


Plots instead of tables?
```{r plot spp change}
levels(shelf_area_change_hemis_reg_stats$region)
#add return to labels
shelf_area_change_hemis_reg_stats[,region:= factor(region, levels = c("East Atlantic", "East Indian"  , "East Pacific",  "West Atlantic", "West Indian" ,  "West Pacific" ), labels = c("East Atlantic", "East Indian"  , "East Pacific",  "West Atlantic", "West Indian" ,  "West Pacific" ))]

(change_area_reg_hemis <- ggplot(data = shelf_area_change_hemis_reg_stats) +
  geom_col(aes(x = region, y = sum_change_hemis_reg/1000, fill = change_direction)) +
  facet_wrap(~hemisphere, nrow = 2, ncol = 1, scales = "free") +
  labs(x = "Region", y = expression(paste(Delta, " Area with 2˚ Latitudinal Shift (1000s of ", km^{2},")"))) +
  scale_fill_viridis_d(option = "A", begin = 0.7, end = 0.3, name = "Direction of Shelf\nArea Change\nWith 2\u00B0 Shift", labels = c("Positive", "Negative")) +
  geom_hline(yintercept = 0) +
  theme_classic())

ggsave(height = 10, width = 3, file = "shelf_area_change_hemis_reg_stats.jpg")

#add return to labels
shelf_spp_change_hemis_reg_stats[,region:= factor(region, levels = c("East Atlantic", "East Indian"  , "East Pacific",  "West Atlantic", "West Indian" ,  "West Pacific" ), labels = c("East Atlantic", "East Indian"  , "East Pacific",  "West Atlantic", "West Indian" ,  "West Pacific" ))]

(change_spp_reg_hemis <- ggplot(data = shelf_spp_change_hemis_reg_stats) +
  geom_col(aes(x = region, y = spp_change_levin_by_reg_hemis, fill = change_direction)) +
  facet_wrap(~hemisphere, nrow = 2, ncol = 1, scales = "free") +
  labs(x = "Region", y =expression(paste(Delta, " in Species Richness with 2˚ Latitudinal Shift"))) +
  scale_fill_viridis_d(option = "A", begin = 0.7, end = 0.3, name = "Change in Richness\nWith 2\u00B0 Shift") +
  geom_hline(yintercept = 0) +
  theme_classic())

ggsave(height = 10, width = 3, file = "shelf_spp_change_hemis_reg_stats.jpg")

library(cowplot)

plot_grid(change_area_reg_hemis, change_spp_reg_hemis, labels = c("a.","b."))

ggsave(height = 4, width = 10, file = "spp_area_change_Levin2009.jpg")

```
What if instead of this type of plot, we do a bar chart ordered by value?
```{r}
#merge two columns
shelf_spp_change_hemis_reg_stats[,Coastline := paste(hemisphere,region,sep = "\n")]

#reorder
levels(shelf_spp_change_hemis_reg_stats$change_direction)
shelf_spp_change_hemis_reg_stats[,change_direction := factor(change_direction, levels = c("Loss","Gain"))]

(ordered_loss_gain_bycoastline <- ggplot(shelf_spp_change_hemis_reg_stats) +
  geom_col(aes(x = reorder(Coastline, spp_change_levin_by_reg_hemis), y = spp_change_levin_by_reg_hemis, fill = change_direction), alpha = 0.8) +
  xlab("Coastline") +
  ylab("Predicted Species Richness\nChange With 2\u00B0 Latitudinal Shift") +
  theme_classic() +
scale_fill_viridis_d(option = "A", begin = 0.3, end = 0.7, name = "Change in Richness\nWith 2\u00B0 Shift") +
  geom_hline(yintercept = 0, size = 0.3) +
  theme(axis.text.x = element_text(angle = 90, size = 10, hjust = 1, vjust = 0.5), legend.title = element_blank()))

save(ordered_loss_gain_bycoastline, file = "ordered_loss_gain_bycoastline.RData")

```



Species Area Curves

How much total area to begin with? #FIX THIS
```{r total area to begin with}
OG_total_shelf_area <- sum(shelf_areas$area_rasterarea)
new_total_shelf_area <- sum(shelf_areas[latitude_start > 0])

final_shelf_area <- OG_total_shelf_area-106656.4

#Equation Spp=16.18*A^0.226 (Phillip Levin 2008) https://besjournals.onlinelibrary.wiley.com/doi/full/10.1111/j.1365-2664.2009.01667.x

OG_species <- 16.18*OG_total_shelf_area^0.226
final_species <- 16.18*final_shelf_area^0.226

final_species -  OG_species

```

Comparing alternative ways of calculating SAR

Also, Chittaro et al. 2010 (Levin is on this paper as well)
spp = 26.1*A^0.183

```{r comparing alternative ways of calculating SAR}
testareas <- data.table(area= seq(1,10000000,by = 100))
testareas[,Levin_SAR := 16.18*area^0.226][,Chittaro_SAR := 26.1*area^0.183][,Meixler_SAR_30 := exp(6.854 + 0.185 * log(area) -1.084 * log(30))][,Meixler_SAR_1 := exp(6.854 + 0.185 * log(area) -1.084 * log(1))][,Meixler_SAR_5 := exp(6.854 + 0.185 * log(area) -1.084 * log(5))][,Meixler_SAR_60 := exp(6.854 + 0.185 * log(area) -1.084 * log(60))][,Meixler_SAR_80 := exp(6.854 + 0.185 * log(area) -1.084 * log(80))][,Meixler_SAR_70 := exp(6.854 + 0.185 * log(area) -1.084 * log(70))][,Meixler_SAR_10 := exp(6.854 + 0.185 * log(area) -1.084 * log(10))] #includes alternative latitudes

#wide to long
testareas.l <- melt(testareas, measure.vars = c(2:ncol(testareas)))

testareas.l$variable <- factor(testareas.l$variable, levels = c("Levin_SAR","Chittaro_SAR","Meixler_SAR_1" , "Meixler_SAR_5"  ,    "Meixler_SAR_10" ,  "Meixler_SAR_30" , "Meixler_SAR_60","Meixler_SAR_70",
"Meixler_SAR_80" ))

levels(testareas.l$variable)

ggplot(data = testareas.l[variable == "Chittaro_SAR" | variable == "Levin_SAR"], aes(x = area)) +
  geom_point(aes(y = value, color = variable)) +
  labs(x = "Area km^2", y = "# spp") +
  theme_classic()

ggsave(file = "testing_SAR.jpg")

```
Continuing with Marci's calculations

spp = exp(6.854 + (0.185 x log(A)) -  (1.084 * log(L)))

Let's make this same figure as area_change_latitude but for species lost or gained instead of area lost or gained

```{r change in # spp}
#Meixler
(spp_change_latitude_meixler  <- ggplot(data = shelf_areas) +
  geom_col(aes(x=latitude_end, y=change_spp_predicted_meixler, fill = spp_change_pos_neg_meixler)) +
 labs(x=paste0("Latitude ","\u00B0","E"), y = "Spp Richness") +
     scale_fill_viridis_d(option = "A", begin = 0.3, end = 0.7, name = "Meixler 2019\n\n Richness with 2\u00B0 Shift") +
  geom_vline(xintercept = 0) +
  geom_hline(yintercept = 0) +
  coord_flip() +
  theme_classic() +
   facet_wrap(~region, scales = "free"))
  

ggsave(spp_change_latitude_meixler, file = "mexlier_spprichness_change_latitude.jpg")

#Meixler ONLY for regions it was calibrated for 

#Meixler
(spp_change_latitude_meixler_reduced  <- ggplot(data = shelf_areas[latitude_end_abs > 22 & latitude_end_abs > 46]) +
  geom_col(aes(x=latitude_end, y=change_spp_predicted_meixler, fill = spp_change_pos_neg_meixler)) +
 labs(x=paste0("Latitude ","\u00B0","E"), y = "Spp Richness") +
     scale_fill_viridis_d(option = "A", begin = 0.3, end = 0.7, name = "Meixler 2019\nChange in Species\n Richness with 2\u00B0 Shift") +
  geom_vline(xintercept = 0) +
  geom_hline(yintercept = 0) +
  coord_flip() +
  theme_classic() +
   facet_wrap(~region, scales = "free") +
    lims(y = c(-120, 55)))
  

ggsave(spp_change_latitude_meixler_reduced, file = "spp_change_latitude_meixler_reduced.jpg")

#Levin

(spp_change_latitude_levin  <- ggplot(data = shelf_areas) +
  geom_col(aes(x=latitude_end, y=change_spp_predicted_levin, fill = spp_change_pos_neg_levin)) +
 labs(x=paste0("Latitude ","\u00B0","E"), y = "Spp Richness") +
     scale_fill_viridis_d(option = "A", begin = 0.3, end = 0.7, name = "Levin 2009\nChange in Species\n Richness with 2\u00B0 Shift") +
  geom_vline(xintercept = 0) +
  geom_hline(yintercept = 0) +
  coord_flip() +
  theme_classic() +
   facet_wrap(~region, scales = "free") +
    lims(y = c(-120, 55)))
  

ggsave(spp_change_latitude_levin, file = "levin_spprichness_change_latitude.jpg")

#Levin (only for limited Meixler latitudes)

(spp_change_latitude_levin_reduced  <- ggplot(data = shelf_areas[latitude_end_abs > 22 & latitude_end_abs > 46]) +
  geom_col(aes(x=latitude_end, y=change_spp_predicted_levin, fill = spp_change_pos_neg_levin)) +
 labs(x=paste0("Latitude ","\u00B0","E"), y = "Spp Richness") +
     scale_fill_viridis_d(option = "A", begin = 0.3, end = 0.7, name = "Levin 2009\nChange in Species\n Richness with 2\u00B0 Shift") +
  geom_vline(xintercept = 0) +
  geom_hline(yintercept = 0) +
  coord_flip() +
  theme_classic() +
   facet_wrap(~region, scales = "free") +
    lims(y = c(-120, 55)))
  

ggsave(spp_change_latitude_levin_reduced, file = "spp_change_latitude_levin_reduced.jpg")

#view with area?
plot_grid(area_change_latitude, spp_change_latitude_levin, nrow = 2, ncol = 1)
```
Try using Phil Levin Indices instead #Spp = 16.18*A^0.226

What if we compare both?
```{r compare Levin versus Meixler}
ggplot(data = shelf_areas) +
  geom_point(aes(x = latitude_end_abs, y = original_spp_predicted_meixler), color = "blue") +
    geom_point(aes(x = latitude_end_abs, y = original_spp_predicted_levin), color = "green") +
  theme_classic()
```

But, what about Chittaro et al. 2010 (Levin is on this paper as well) (Honestly can't figure why #s are different, let's go with newer is better?)

spp = 26.1*A^0.183
```{r}

```

