---
title: "R Notebook"
output: html_notebook
---

Test with a single region

---
title: "Starting with Bathy for Shelves"
output: html_notebook
---

```{r setup}
library(raster)
library(rgdal)
library(sf)
library(ncdf4)
library(rmapshaper)
library(tidyverse)
library(diptest)
library(moments)
library(viridis) #colors
library(data.table)
library(hydroTSM) #hypsometric curves
library(gridExtra)
library(ggfortify)
library(rgeos)
library(ggridges) #allows for gradient in density plot
library(psych) #summary statistics
library(plyr) #round_any

etopo_shelf_df <- readRDS("~/Documents/grad school/Rutgers/Repositories/shelf_habitat_distribution/etopo_shelf_df.rds")
#bring in bathy for shelf regions

#LMEs
LME_spdf <- readOGR("LME66/LMEs66.shp")


#get rid of Antarctica and the arctic
LME_spdf <- LME_spdf[LME_spdf$LME_NUMBER != 61 & LME_spdf$LME_NUMBER != 64,]

#equal area projection
eaproj <- CRS("+proj=laea")
LME_spdf.EA <- spTransform(LME_spdf, eaproj)


```


Let's just start with one LME to get started
```{r LME_baltic}
LME_baltic <- LME_spdf[1,]
plot(LME_baltic)

center <- gCentroid(LME_baltic)
center@coords[2]
```


Make bathy layer into raster (I should see if this is different if I use equal area projection (To Do))
```{r rasterize}
proj <- crs(LME_spdf)
etopo_shelf_raster <- rasterFromXYZ(etopo_shelf_df, crs = proj)


#make figure of shelves
  #png("etopo_shelf_bathy.png", height=5, width=7, units="in", res=300)
  #plot(etopo_shelf_raster)
  #dev.off()

```

Depth distribution
```{r depth distribution}
mean(etopo_shelf_raster$z)
hist(etopo_shelf_raster$z)

cellStats(etopo_shelf_raster, stat = 'mean')
hist(etopo_shelf_raster)
```

Clip raster to each individual polygon

```{r clip raster to polygon}
#crop bathy layer to LME subset
bathy_baltic_extent <- crop(etopo_shelf_raster, extent(LME_baltic))

#which areas of raster fall within borders?
bathy_baltic_mask <- mask(bathy_baltic_extent, LME_baltic)

#convert to df for plotting

  # First, to a SpatialPointsDataFrame
  bathy_baltic_pts <- rasterToPoints(bathy_baltic_mask, spatial = TRUE)
  # Then to a 'conventional' dataframe
  bathy_baltic_dt  <- data.table(data.frame(bathy_baltic_pts))

#keep depth positive  
bathy_baltic_dt[,depth := -(z)]

#fortify to plot shapefile on top
#LME_baltic_df <- fortify(LME_baltic) (don't need because indstead I make polygon to surround raster)

#To get a polygon that surrounds cells that are not NA

  # make all values the same.
  bathy_baltic_mask_1s <- reclassify(bathy_baltic_mask, cbind(-Inf, Inf, 1))
  
  # convert to polygons (you need to have package 'rgeos' installed for this to work)
  bathy_baltic_polygon_1s <- rasterToPolygons(bathy_baltic_mask_1s, dissolve=TRUE) #single polygon
  bathy_baltic_polygon <- rasterToPolygons(bathy_baltic_mask, dissolve=TRUE) #one polygon for each depth profile
  
  LME_reduced_baltic_df <- fortify(bathy_baltic_polygon)

ggplot(bathy_baltic_dt) +
  geom_tile(aes(x = x, y = y, fill = depth)) +
  scale_fill_viridis(direction = -1, limits = c(0,2000)) +
  theme_classic() +
  labs(x = "Longitude", y = "Latitude", fill = "Depth (m)") +
  guides(fill = guide_colorbar(reverse = T)) +
  geom_path(data = LME_reduced_baltic_df, aes(x = long, y = lat, group = group), fill = NA, color = "black")

summary(bathy_baltic_dt)

#how to get area of raster

  #get sizes of all cells in raster [km2]
  cell_size<-area(bathy_baltic_mask, na.rm=TRUE, weights=FALSE)
  #delete NAs from vector of all raster cells
  ##NAs lie outside of the rastered region, can thus be omitted
  cell_size<-cell_size[!is.na(cell_size)]
  #compute area [km2] of all cells in geo_raster
  bathy_raster_area<-length(cell_size)*median(cell_size)
  
#Also, try using raster::area on polygon

#reproject
#The Lambert azimuthal equal-area projection is a particular mapping from a sphere to a disk. It accurately represents area in all regions of the sphere, but it does not accurately represent angles.
  equalareaprojection<- crs("+proj=laea")

  bathy_baltic_polygon_1s.EA <- spTransform(bathy_baltic_polygon_1s, equalareaprojection)
  bathy_baltic_polygon.EA <- spTransform(bathy_baltic_polygon, equalareaprojection)
  
    #get sizes of all polygons in shapefile (m^2)
  polygon_area <- area(bathy_baltic_polygon.EA, dissolve = T)

  polygon_area_kms <- polygon_area/1e6

```

Will have to be creative on how to map LMEs that cross 180 longitude, ex. Aleutian Islands
```{r high longitude examples}
LME_datatable <- LME_spdf@data
LME_aleutian <- LME_spdf[LME_spdf$LME_NUMBER == 65,]

#crop bathy layer to LME subset
bathy_aleutian_extent <- crop(etopo_shelf_raster, extent(LME_aleutian))

#which areas of raster fall within borders?
bathy_aleutian_mask <- mask(bathy_aleutian_extent, LME_aleutian)

#convert to df for plotting

  # First, to a SpatialPointsDataFrame
  bathy_aleutian_pts <- rasterToPoints(bathy_aleutian_mask, spatial = TRUE)
  # Then to a 'conventional' dataframe
  bathy_aleutian_dt  <- data.table(data.frame(bathy_aleutian_pts))

#keep depth positive  
bathy_aleutian_dt[,depth := -(z)]

bathy_aleutian_dt$x_shift <- ifelse(bathy_aleutian_dt$x >0, bathy_aleutian_dt$x - 360, bathy_aleutian_dt$x)


#To get a polygon that surrounds cells that are not NA

  # make all values the same.
  r <- reclassify(bathy_aleutian_mask, cbind(-Inf, Inf, 1))
  
  # convert to polygons (you need to have package 'rgeos' installed for this to work)
  pp <- rasterToPolygons(r, dissolve=TRUE)
  
  LME_reduced_aleutian_df <- fortify(pp)
  
  
LME_reduced_aleutian_df$long_shift <- ifelse(LME_reduced_aleutian_df$long >0, LME_reduced_aleutian_df$long - 360, LME_reduced_aleutian_df$long)
  

ggplot(bathy_aleutian_dt) +
  geom_tile(aes(x = x_shift, y = y, fill = depth)) +
  scale_fill_viridis(direction = -1, limits = c(0,2000)) +
  theme_classic() +
  labs(x = "Longitude (˚E)", y = "Latitude", fill = "Depth (m)") +
  guides(fill = guide_colorbar(reverse = T)) +
  geom_path(data = LME_reduced_aleutian_df, aes(x = long_shift, y = lat, group = group), fill = NA, color = "black")

```

How does area vary with depth
```{r hyposometric curves}
#list of values within raster
bathy_baltic_depth_list <- getValues(bathy_baltic_mask)
bathy_baltic_depth_list <- bathy_baltic_depth_list[!is.na(bathy_baltic_depth_list)]

bathy_baltic_depth_list_pos <- -1*bathy_baltic_depth_list

bathy_baltic_SGDF <- as(bathy_baltic_mask, 'SpatialGridDataFrame') # as spatial grid data frame
hypsometric(bathy_baltic_SGDF)

#dimensions of raster?
dim(bathy_baltic_mask)

#total number of cells?
cells <- dim(bathy_baltic_mask)[1]*dim(bathy_baltic_mask)[2]


#cells of each depth value in data table
bathy_baltic_bydepth <- data.table(freq(bathy_baltic_mask))

#NA cells
na_cell_count <- bathy_baltic_bydepth[is.na(value),]$count

#filled cells

cells_filled <- cells-na_cell_count

bathy_baltic_bydepth_complete <- bathy_baltic_bydepth[!is.na(value)]

#calculate percent of total area in polygon
bathy_baltic_bydepth_complete[,percent_area := count/sum(count)]


bathy_baltic_bydepth_complete[,area := percent_area*bathy_raster_area][,depth := -value]



(plot <- ggplot(data = bathy_baltic_bydepth_complete, aes(x = depth, y = area, fill = depth)) +
  geom_col(width = 15) +
  theme_classic() +
  labs(x = "Depth (m)", y = "Area (km^2)", fill = "Depth (m)") +
  scale_y_continuous(expand = c(0, 0)) +
  scale_x_continuous(expand = c(0,0)) +
  scale_fill_gradientn(colors = rev(viridis(10)), limits = c(0, 2000)) +
  guides(position = "bottom", fill = guide_colorbar(reverse = T)))
  

ggsave(plot, filename = "Baltic_Sea.png")
```

I now need a datatable with:

depth bin              area         spp # 
0-15, 16-30, etc. 

TO DO: LOOK INTO WHAT COUNT MEANS
```{r making depth bin table }
bathy_baltic_bydepth_complete[, depth_15s := round_any(depth, 15, f = floor)] #grouping to 15 m depths

#sum areas over depth bins
bathy_baltic_bydepth_complete[, area_summed_15mbins :=  sum(area), depth_15s]

bathy_baltic_bydepth_complete_15mbins <- unique(bathy_baltic_bydepth_complete[,.(depth_15s, area_summed_15mbins)])

#make sure it looks right
ggplot(data = bathy_baltic_bydepth_complete_15mbins, aes(x = depth_15s, y = area_summed_15mbins)) + geom_col() +
  theme_classic()
```
TO DO: check if we're actually in km and not meters (yes, looks good)

Calculate # spp using Levin et al. 2009
```{r }
bathy_baltic_bydepth_complete_15mbins[,original_spp_predicted_levin := 16.18*area_summed_15mbins^0.226]

#spp by depth plot
ggplot(data = bathy_baltic_bydepth_complete_15mbins, aes(x = depth_15s)) +
  geom_col(aes(y = area_summed_15mbins/100)) +
  geom_point(aes(y = original_spp_predicted_levin)) +
  labs(x = "Depth  (15 m bins)", y = "Area (100s of km^2)") +
  theme_classic()

#be sure to order table so that shift is used accurately (I did this above also, this is just a check)
#include absolute value of latitude end
setkey(bathy_baltic_bydepth_complete_15mbins, depth_15s)

bathy_baltic_bydepth_complete_15mbins[, change_spp_predicted_levin :=  original_spp_predicted_levin -  shift(original_spp_predicted_levin, n = 1, type = "lag")] #change in species from previous bin to this bin?

bathy_baltic_bydepth_complete_15mbins[, spp_change_pos_neg_levin := ifelse(change_spp_predicted_levin > 0, "Positive","Negative")] #for coloring


#overall sum of spp richness change
total_spp_richness_change_baltic <- sum(bathy_baltic_bydepth_complete_15mbins$change_spp_predicted_levin, na.rm = T)


#plot change in spp
#spp by depth plot
ggplot(data = bathy_baltic_bydepth_complete_15mbins, aes(x = depth_15s)) +
  geom_col(aes(y = change_spp_predicted_levin, fill = spp_change_pos_neg_levin)) +
   scale_fill_viridis_d(option = "A", begin = 0.3, end = 0.7, name = paste0("Overall Species\nRichness Change: ", round(total_spp_richness_change_baltic,2), "\n\nDirection of Species\nRichness Change with\n15m Depth Shift"), labels = c("Negative", "Positive")) +
  labs(x = "Depth  (15 m bins)", y = "Change in Spp Richness") +
  geom_hline(yintercept = 0) +
  theme_classic()

```
Now I have to create this for each region in a loop, and then export to PDF

```{r loop for spp loss or gain calculations}

#reorder LME_spdf by LME # 
LME_spdf_ordered <- LME_spdf[order(LME_spdf$LME_NUMBER),]


LME_area_spp_change_stats <- as.data.table(matrix(nrow = nrow(LME_spdf)))
LME_area_spp_change_stats[, lme_name := as.factor(V1)][, lme_number := as.numeric(V1)][,lme_area := as.numeric(V1)][, total_area_change := as.numeric(V1)][, total_spp_change := as.numeric(V1)]

LME_area_spp_change_stats[, V1 := NULL]


lme_spp_change_plots <- vector("list", length = nrow(LME_spdf))
lme_spp_count_area_sum_plots <- vector("list", length = nrow(LME_spdf))

for (i in 1:nrow(LME_area_spp_change_stats)) {
  LME_single <- LME_spdf_ordered[i,]
  LME_area_spp_change_stats[i, "lme_name"] <- LME_single$LME_NAME
  LME_area_spp_change_stats[i, "lme_number"] <- LME_single$LME_NUMBER
  
  #clip raster to LME
  LME_extent <- crop(etopo_shelf_raster, extent(LME_single))
  
  #which areas of raster fall within borders?
  LME_mask <- mask(LME_extent, LME_single)
  
  #conver to df for plotting
    # First, to a SpatialPointsDataFrame
    LME_points <- rasterToPoints(LME_mask, spatial = TRUE)
    
    # Then to a 'conventional' dataframe
    LME_dt  <- data.table(data.frame(LME_points))
  
  #keep depth positive  
  LME_dt[,depth := -(z)]
  
  LME_dt$x_shift <- ifelse(
    max(LME_dt$x)-min(LME_dt$x) > 200 & LME_dt$x >0,
    LME_dt$x - 360,
    LME_dt$x)
  
  #size of raster  
  #get sizes of all cells in raster [km2]
    cell_size<-area(LME_mask, na.rm=TRUE, weights=FALSE)
    #delete NAs from vector of all raster cells
    ##NAs lie outside of the rastered region, can thus be omitted
    cell_size<-cell_size[!is.na(cell_size)]
    #compute area [km2] of all cells in geo_raster
    bathy_raster_area<-length(cell_size)*median(cell_size)
    
    LME_area_spp_change_stats[i, "lme_area"] <- bathy_raster_area
  
  
    #list of values within raster
  LME_bathy_depth_list <- getValues(LME_mask)
  LME_bathy_depth_list <- LME_bathy_depth_list[!is.na(LME_bathy_depth_list)]
  
  LME_bathy_depth_list_pos <- -1*LME_bathy_depth_list


  #cells of each depth value in data table
  LME_bathy_bydepth <- data.table(freq(LME_mask))
  
  LME_bathy_bydepth_complete <- LME_bathy_bydepth[!is.na(value)]
  
  #calculate percent of total area in polygon
  LME_bathy_bydepth_complete[,percent_area := count/sum(count)]
  
  
  LME_bathy_bydepth_complete[,area := percent_area*bathy_raster_area][,depth := -value]
  
  LME_bathy_bydepth_complete[, depth_15s := round_any(depth, 15, f = floor)] #grouping to 15 m depths

  #sum areas over depth bins
  LME_bathy_bydepth_complete[, area_summed_15mbins :=  sum(area), depth_15s]

  LME_bathy_bydepth_complete_15mbins <- unique(LME_bathy_bydepth_complete[,.(depth_15s, area_summed_15mbins)])
  
  LME_bathy_bydepth_complete_15mbins[,original_spp_predicted_levin := 16.18*area_summed_15mbins^0.226]

  #spp by depth plot
  lme_spp_count_area_sum_plots[[i]] <- ggplot(data = LME_bathy_bydepth_complete_15mbins, aes(x = depth_15s)) +
    geom_col(aes(y = area_summed_15mbins/100)) +
    geom_point(aes(y = original_spp_predicted_levin)) +
    labs(x = "Depth  (15 m bins)", y = "Area (100s of km^2)") +
    theme_classic()

  #be sure to order table so that shift is used accurately (I did this above also, this is just a check)
  #include absolute value of latitude end
  setkey(LME_bathy_bydepth_complete_15mbins, depth_15s)
  
  #change in area
    LME_bathy_bydepth_complete_15mbins[, change_area :=  area_summed_15mbins -  data.table::shift(area_summed_15mbins, n = 1, type = "lag")] #change in species from previous bin to this bin?
  #change in spp richness
  LME_bathy_bydepth_complete_15mbins[, change_spp_predicted_levin :=  original_spp_predicted_levin -  data.table::shift(original_spp_predicted_levin, n = 1, type = "lag")] #change in species from previous bin to this bin?

  LME_bathy_bydepth_complete_15mbins[, spp_change_pos_neg_levin := ifelse(change_spp_predicted_levin > 0, "Positive","Negative")] #for coloring

    #overall change in area
  LME_area_spp_change_stats[i, "total_area_change"] <- sum(LME_bathy_bydepth_complete_15mbins$change_area, na.rm = T)
  
  #overall sum of spp richness change
  LME_area_spp_change_stats[i, "total_spp_change"] <- sum(LME_bathy_bydepth_complete_15mbins$change_spp_predicted_levin, na.rm = T)


  #plot change in spp
  #spp by depth plot
  lme_spp_count_area_sum_plots[[i]] <- ggplot(data = LME_bathy_bydepth_complete_15mbins, aes(x = depth_15s)) +
    geom_col(aes(y = change_spp_predicted_levin, fill=spp_change_pos_neg_levin), alpha = 0.5) +
    scale_fill_viridis_d(option = "A", begin = 0.3, end = 0.7, name = paste0("LME: ", LME_area_spp_change_stats[i, "lme_number"],"\n", LME_area_spp_change_stats[[i, "lme_name"]],"\n\nSpecies Richness Change: ", round(LME_area_spp_change_stats[i, "total_spp_change"],2), "\n\nSpecies Richness Change\nwith 15m Depth Shift"), breaks = c("Positive","Negative"), labels = c("Gain", "Loss")) +
    labs(x = "Depth  (15 m bins)", y = expression(Delta~'Species Richness')) +
#    geom_hline(yintercept = 0) +
    theme_classic() +
    scale_x_continuous(expand = c(0,0)) +
    scale_y_continuous(expand = c(0,0)) +
    theme(legend.title = element_text(size = 4), 
               legend.text = element_text(size = 4), 
                legend.key.size = unit(.5,'cm'))

 
    print(paste(i, nrow(LME_area_spp_change_stats), sep = "/"))
  

}

saveRDS(LME_area_spp_change_stats, file = "LME_area_spp_change_stats.rds")
saveRDS(lme_spp_count_area_sum_plots, file = "lme_spp_count_area_sum_plots.rds")

#plots to pdf
ggsave("arranged_LME_sppchange.pdf", marrangeGrob(lme_spp_count_area_sum_plots, ncol=2, nrow = 2))

```
Dummy legend
```{r}
dummy_legend_depth_richness_plot <- ggplot(data = LME_bathy_bydepth_complete_15mbins, aes(x = depth_15s)) +
    geom_col(aes(y = change_spp_predicted_levin, fill=spp_change_pos_neg_levin), alpha = 0.5) +
    scale_fill_viridis_d(option = "A", begin = 0.3, end = 0.7, breaks = c("Positive","Negative"), labels = c("Gain", "Loss")) +
    labs(x = "Depth  (15 m bins)", y = expression(Delta~'Species Richness')) +
    theme_classic() +
    scale_x_continuous(expand = c(0,0)) +
    scale_y_continuous(expand = c(0,0)) +
    theme(legend.title = element_text(size = 4), 
               legend.text = element_text(size = 4), 
                legend.key.size = unit(.5,'cm'))

saveRDS(dummy_legend_depth_richness_plot, here::here("Figures","Figure1","dummy_legend_richness_plot.rds"))
```


Instead of change in spp richness, I will just try plotting overall richness for each 15 m depth bin
```{r richness each 15 m depth bin}

#reorder LME_spdf by LME # 
LME_spdf_ordered <- LME_spdf[order(LME_spdf$LME_NUMBER),]


LME_area_spp_richness_stats <- as.data.table(matrix(nrow = nrow(LME_spdf)))
LME_area_spp_richness_stats[, lme_name := as.factor(V1)][, lme_number := as.numeric(V1)][,lme_area := as.numeric(V1)][, total_area := as.numeric(V1)][, total_spp_richness := as.numeric(V1)]

LME_area_spp_richness_stats[, V1 := NULL]

lme_spp_richness_plots <- vector("list", length = nrow(LME_spdf))

for (i in 1:nrow(LME_area_spp_richness_stats)) {
  LME_single <- LME_spdf_ordered[i,]
  LME_area_spp_richness_stats[i, "lme_name"] <- LME_single$LME_NAME
  LME_area_spp_richness_stats[i, "lme_number"] <- LME_single$LME_NUMBER
  
  #clip raster to LME
  LME_extent <- crop(etopo_shelf_raster, extent(LME_single))
  
  #which areas of raster fall within borders?
  LME_mask <- mask(LME_extent, LME_single)
  
  #conver to df for plotting
    # First, to a SpatialPointsDataFrame
    LME_points <- rasterToPoints(LME_mask, spatial = TRUE)
    
    # Then to a 'conventional' dataframe
    LME_dt  <- data.table(data.frame(LME_points))
  
  #keep depth positive  
  LME_dt[,depth := -(z)]
  
  LME_dt$x_shift <- ifelse(
    max(LME_dt$x)-min(LME_dt$x) > 200 & LME_dt$x >0,
    LME_dt$x - 360,
    LME_dt$x)
  
  #size of raster  
  #get sizes of all cells in raster [km2]
    cell_size<-area(LME_mask, na.rm=TRUE, weights=FALSE)
    #delete NAs from vector of all raster cells
    ##NAs lie outside of the rastered region, can thus be omitted
    cell_size<-cell_size[!is.na(cell_size)]
    #compute area [km2] of all cells in geo_raster
    bathy_raster_area<-length(cell_size)*median(cell_size)
    
    LME_area_spp_richness_stats[i, "lme_area"] <- bathy_raster_area
  
  
    #list of values within raster
  LME_bathy_depth_list <- getValues(LME_mask)
  LME_bathy_depth_list <- LME_bathy_depth_list[!is.na(LME_bathy_depth_list)]
  
  LME_bathy_depth_list_pos <- -1*LME_bathy_depth_list


  #cells of each depth value in data table
  LME_bathy_bydepth <- data.table(freq(LME_mask))
  
  LME_bathy_bydepth_complete <- LME_bathy_bydepth[!is.na(value)]
  
  #calculate percent of total area in polygon
  LME_bathy_bydepth_complete[,percent_area := count/sum(count)]
  
  
  LME_bathy_bydepth_complete[,area := percent_area*bathy_raster_area][,depth := -value]
  
  LME_bathy_bydepth_complete[, depth_15s := round_any(depth, 15, f = floor)] #grouping to 15 m depths

  #sum areas over depth bins
  LME_bathy_bydepth_complete[, area_summed_15mbins :=  sum(area), depth_15s]

  LME_bathy_bydepth_complete_15mbins <- unique(LME_bathy_bydepth_complete[,.(depth_15s, area_summed_15mbins)])
  
  LME_bathy_bydepth_complete_15mbins[,original_spp_predicted_levin := 16.18*area_summed_15mbins^0.226]


  #plot change in spp
  #spp by depth plot
  lme_spp_richness_plots[[i]] <- ggplot(data = LME_bathy_bydepth_complete_15mbins, aes(x = depth_15s)) +
    geom_area(aes(y = original_spp_predicted_levin), fill= "bisque") +
    labs(x = "Depth (15 m bins)", y = "Species Richness", title = LME_single$LME_NAME) +
    scale_x_continuous(expand=c(0,0)) +
    scale_y_continuous(expand=c(0,0)) +
    theme_classic() +
    theme(legend.title = element_text(size = 4), 
               legend.text = element_text(size = 4), 
                legend.key.size = unit(.5,'cm'))

 
    print(paste(i, nrow(LME_area_spp_richness_stats), sep = "/"))
  

}

saveRDS(LME_area_spp_richness_stats, file = "LME_area_spp_richness_stats.RData")

#plots to pdf
ggsave("arranged_LME_spp_richness.pdf", marrangeGrob(lme_spp_richness_plots, ncol=2, nrow = 2))
```
```


Add column for percent loss of starting area
```{r}
LME_area_spp_change_stats[,percent_change := total_area_change/lme_area]
```

