---
title: "Depth_Shelf"
author: "Professor Becca Selden edited by Zoë Kitchel"
date: "5/20/2020; 10/27/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(raster)
library(rgdal)
library(sf)
library(ncdf4)
library(rmapshaper)
library(tidyverse)
```

### Read in ETOPO1
Source: https://www.ngdc.noaa.gov/mgg/global/relief/ETOPO1/data/bedrock/cell_registered/netcdf/
Even on my computer, this was just such a huge file that many of the operations I tried below were SUPER slow, so I think I am advocating for keeping ETOPO2 for now, and see if reviewers want the higher resolution. Zoë agrees.
```{r etopo1, eval=F}
#etopo <- raster("ETOPO1_Bed_f_gdal.grd"), 
#                crs="+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0 ")

```

### Read in ETOPO2
```{r etopo2}
etopo2 <- raster("DepthShift2.0/Data/ETOPO2v2g_f4.nc")
```

### Restrict to depths less than 2000
```{r etopo less than 2000}
etopo2[etopo2 >=0] <- NA
etopo2[etopo2< -2000] <- NA
png("DepthShift2.0/Figures/bathy_2000m.png", height=5, width=7, units="in", res=300)
plot(etopo2)
dev.off()
```

### Convert etopo to dataframe
```{r etopo df}
#etopo_df <- as.data.frame(etopo2, xy=TRUE)
```

### Read in Continental Shelf Data
Source:https://www.bluehabitats.org/?page_id=58
```{r shelf}
shelf <- readOGR("Shelf_BlueHabitats/Shelf.shp")
#readOGR: rgdal: reads OGR vector maps into spatial objects

shelf_tv <- read_sf("Shelf_BlueHabitats/Shelf.shp")
#read_sf read simple features from file or database, or retrieve layer names and their geometry types (essentially metadata)


shelf_simp <- ms_simplify(shelf_tv, keep=0.05) #keep 5% of the vertices
#uses mapshaper to simplify polygons

shelf_buffer <- st_buffer(shelf_simp, dist=0.3, endCapStyle="ROUND")
#computes a buffer around each geometry, 0.3 = buffer distance


ggplot(shelf_simp) + 
  geom_sf(fill="black", col="gray") +
  theme_bw()
```



### Crop ETOPO to continental shelf
#this takes a long time, so I'll just bring back in Becca's final product
```{r etopo crop}
etopo_shelf <- mask(etopo2, shelf_simp)


writeRaster(etopo_shelf, here::here("Data/etoposhelf.grd"))

etopo_df <- as.data.frame(etopo2, xy=T)


etopo_shelf_df <- as.data.frame(etopo_shelf, xy=T)

etopo_shelf_lim <- subset(etopo_shelf_df, !(is.na(layer)))



png(here::here("Figures/hist_shelfdepth.png"), height=8, width=5, units="in", res=300)
par(mfrow=c(2,1))
hist(etopo_df$layer, main="ETOPO less than 2000m")
hist(etopo_shelf_lim$layer, main="ETOPO masked by shelf polygons")
dev.off()

png(here::here("Figures/shelfpts.png"), height=5, width=7, units="in", res=300)
plot(y ~ x, etopo_shelf_lim, cex=0.01)
dev.off()

saveRDS(etopo_shelf_lim, here::here("Data/etopo_shelf_lim.rds"))
saveRDS(etopo_shelf_df, here::here("Data/etopo_shelf_df.rds"))
```

### Convert back to raster
```{r rasterize}

etopo_shelf_raster <- rasterFromXYZ(etopo_shelf_df)
png(here::here("Figures/etopo_shelf_bathy.png"), height=5, width=7, units="in", res=300)
plot(etopo_shelf_raster)
dev.off()

etopo_shelf_lim_raster <- rasterFromXYZ(etopo_shelf_lim)
plot(etopo_shelf_lim_raster)
```
