---
title: "R Notebook"
output: html_notebook
---

Here, I am developing a script to post onto stack exchange to get a better sense of how to calculate area correctly. 

```{r setup}
library(rgeos)
library(raster)
library(sp)
library(rgdal) #readOGR
```

Pull in greenland shapefile
```{r play shapefile}

outer <- Polygon(rbind(c(-15.5376528679349, 79.4881322107174), c(-7.48352810724904, 79.908837268938), c(-6.66906605279766, 78.6475974600555), c(-15.4471570841069, 78.7892258712421), c(-15.5376528679349, 79.4881322107174)))

inner <- Polygon(rbind(c(-13.6190362010361, 79.3320228496385), c(-13.7164252183978,79.039680481742), c(-9.04175238503622, 79.039680481742), c(-9.23653041975962, 79.3859727584367), c(-13.6190362010361, 79.3320228496385)), hole = T)

# create lists of polygon objects from polygon objects and unique ID A
# `Polygons` is like a single observation.
h1 <- Polygons(list(outer, inner), "donut1")

# create spatial polygons object from lists A SpatialPolygons is like a
# shapefile or layer.
onedonut <- SpatialPolygons(list(h1))


```

Equal area projection
```{r eaprojection}
projection(onedonut) <- CRS("+proj=eqearth")
```


Convert to raster
```{r to raster}
r <- raster(ncol=100, nrow=100)
extent(r) <- extent(onedonut)
onedonut.raster <- rasterize(onedonut, r)
```

Calculate area using raster::area on non-projected raster
```{r area non projected raster}
   #raster area calculation
      #get sizes of all cells in raster [km2]
    cell_size_raster<-area(onedonut.raster, na.rm=TRUE, weights=FALSE)
    #delete NAs from vector of all raster cells
    ##NAs lie outside of the rastered region, can thus be omitted
    cell_size_raster<-cell_size_raster[!is.na(onedonut.raster)]
    #compute area [km2] of all cells in geo_raster
    segment_area_raster <-length(cell_size_raster)*median(cell_size_raster)/1e6
```

Calculating area using raster::area on projected shapefile
```{r area projected sp}
    cell_size_sp<-area(onedonut, na.rm=TRUE, weights=FALSE)

    #compute area [km2] of all polygons in shape file
    segment_area_equalarea <-length(cell_size_sp)*median(cell_size_sp)/1e6
```

Calculate area using rgeos::gArea on projected shapefile

```{r area projected sp rgeos}
rgeos_shapefile_area <- gArea(onedonut) / 1e6
  
  
```

Are these values close?
```{r}
segment_area_raster
segment_area_equalarea
rgeos_shapefile_area
```

